
\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs,longtable,array}
\usepackage{enumitem}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\usepackage{bookmark}

\setlength{\parskip}{0.45em}
\setlength{\parindent}{0pt}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}
\setlength{\LTleft}{0pt}
\setlength{\LTright}{0pt}
\sloppy
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\renewcommand{\qedsymbol}{$\square$}
\newcommand{\RP}{\mathsf{RP}}
\newcommand{\RC}{\mathsf{RC}}

\makeatletter
\renewcommand{\verbatim@font}{\normalfont\ttfamily\small}
\makeatother


\title{Finite Row-Graded Magmas over \(\mathbb F_3\): Intrinsic Selection and Linearization}
\author{Burundai Taryi\thanks{Independent researcher, \href{mailto:burundai.taryi@gmail.com}{burundai.taryi@gmail.com}.}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study a finite family of \(\sigma\)-equivariant row-graded magmas over \(\mathbb F_3\), with carrier \[
M=S\times S,\qquad S=\mathbb F_3,
\] left-row output, and a fixed same-row Steiner rule. The normal
form is \[
\Omega'\cong S^{18}\times S^3\cong S^{21},\qquad |\Omega'|=3^{21}.
\]

The paper has three layers. First, it maps the finite landscape: the
certified associativity atlas gives \[
63\le \operatorname{Assoc}\le 597,
\] with endpoint loci and exact partition function, while the global
\(\mathcal H\)-frontier theorem gives \[
\min H_{tot}=-2268,\qquad \max H_{tot}=7302,
\qquad \max_{N_-=0}H_{tot}=7020.
\] The layer above the pure frontier consists of twelve impure
signed-cancellation maximizers.

Second, the central selection theorem isolates one magma by four
internal finite conditions: \[
\boxed{
\mathcal H_{acc}=0,
\qquad
H_{diag}=0,
\qquad
\text{row-anchor idempotence},
\qquad
\text{pure }C/J.
}
\] These conditions force the right-row projection magma \(\RP\), whose multiplication is \[
(r_1,c_1)\cdot(r_2,c_2)=
\begin{cases}
(r_1,r_2),&r_1\ne r_2,\\[1mm]
(r_1,\overline{c_1c_2}),&r_1=r_2,
\ c_1\ne c_2,\\[1mm]
(r_1,r_1),&r_1=r_2,
\ c_1=c_2.
\end{cases}
\]

Third, after selection, \(\RP\) admits a linear/operator passport:
rank-three left translations, a finite composition envelope,
endomorphism and intertwiner structure, a binary term algebra of size \[
|T_2(\RP)|=630,
\] a row-contraction matrix bridge, a hidden-continuation operator
identity, and companion carrier-level Lie and signed symplectic
structures.

The organizing implication is \[
\boxed{\text{finite landscape}}
\Longrightarrow
\boxed{\text{intrinsic finite selection}}
\Longrightarrow
\boxed{\text{linearization of the selected magma}}.
\]
\end{abstract}

\tableofcontents
\newpage

\setcounter{section}{-1}
\section{Introduction}\label{introduction}

\subsection{Object and aim}\label{aim}

The object of the paper is a finite landscape of row-graded
multiplications on
\[
M=S\times S,\qquad S=\mathbb F_3.
\]
The first coordinate is the row and the second coordinate is the column.
The family is defined by left-row output \(\mathrm{AX0}\), a fixed
same-row off-diagonal Steiner rule \(\mathrm{AX2}\), and equivariance for
\[
\sigma(r,c)=(r+1,c+1).
\]
These constraints leave exactly twenty-one ternary coordinates,
\[
\Omega'\cong S^{18}\times S^3\cong S^{21},
\qquad |\Omega'|=3^{21}=10\,460\,353\,203.
\]

The paper follows one spine:
\[
\boxed{\text{finite row-graded landscape}}
\quad\Longrightarrow\quad
\boxed{\text{intrinsic finite selection}}
\quad\Longrightarrow\quad
\boxed{\text{linearization after selection}}.
\]
The global atlases describe the ambient landscape.  The compact selector
isolates one point of this landscape.  The operator, term-function,
matrix-carrier, Lie, and symplectic structures then describe the selected
magma.

\subsection{Main theorem}\label{main-theorem}

\begin{theorem}[Main theorem package]\label{thm:main}
Let \(\Omega'\) be the row-graded landscape on \(M=S\times S\), with
left-row output, the fixed same-row off-diagonal Steiner rule, and
\(\sigma\)-equivariance for \(\sigma(r,c)=(r+1,c+1)\).  Then the following
hold.

\begin{enumerate}[label=\textup{(\roman*)}]
\item The landscape has normal form
\[
\Omega'\cong S^{18}\times S^3\cong S^{21}.
\]

\item The finite associativity atlas satisfies
\[
63\le \operatorname{Assoc}\le 597,
\]
with endpoint loci and exact partition function.  The global
\(\mathcal H\)-frontier satisfies
\[
\min H_{tot}=-2268,
\qquad
\max H_{tot}=7302,
\qquad
\max_{N_-=0}H_{tot}=7020.
\]
The layer above the pure frontier consists of twelve impure
signed-cancellation maximizers.

\item The four internal conditions
\[
\mathcal H_{acc}=0,
\qquad
H_{diag}=0,
\qquad
\text{row-anchor idempotence},
\qquad
\text{pure }C/J
\]
select a unique magma in \(\Omega'\).  This selected magma is the
right-row projection magma \(\RP\), with multiplication
\[
(r_1,c_1)\cdot(r_2,c_2)=
\begin{cases}
(r_1,r_2), & r_1\ne r_2,\\[1mm]
(r_1,\overline{c_1c_2}), & r_1=r_2,\ c_1\ne c_2,\\[1mm]
(r_1,r_1), & r_1=r_2,\ c_1=c_2.
\end{cases}
\]

\item After selection, \(\RP\) has a finite linear/operator passport:
every left translation has rank \(3\), the positive composition envelope
has \(51\) maps, the endomorphism monoid consists of the six simultaneous
symbol permutations and the three diagonal collapses, and the binary
term-function algebra has
\[
T_2(\RP)\cong \{x,y\}\times\mathcal S_{21}\times\mathcal U_{15},
\qquad |T_2(\RP)|=630.
\]
The carrier linearization
\[
k[M]\cong \operatorname{Mat}_3(k),
\qquad e_{(r,c)}\leftrightarrow E_{rc},
\]
supports a row-contraction bridge, a hidden-continuation operator identity,
carrier-level compact matrix companions, and a signed symplectic lift of the
pure \(C/J\) directed-edge geometry.

\item The proof package is finite.  The normal form, information selectors,
diagonal compression, and compact selection theorem are hand finite
arguments.  The associativity and post-selection passport data are checked
against finite tables.  The global \(\mathcal H\)-frontier is certified
by a separate finite frontier bundle.
\end{enumerate}
\end{theorem}

\begin{proof}
This theorem is the conjunction of the main results proved below.  The
normal form is Theorem~\ref{thm:1.1}.  The associativity atlas is
Theorem~\ref{thm:5.2}, with endpoint and partition data in
Theorems~\ref{thm:6.1} and~\ref{thm:6.2}.  The \(\mathcal H\)-frontier
is Theorem~\ref{thm:7.2}.  The compact selection theorem is
Theorem~\ref{thm:12.3}.  The operator and term-function statements are
Lemma~\ref{lem:13.1}, Theorems~\ref{thm:14.2},~\ref{thm:15.2},
\ref{thm:16.2}, and~\ref{thm:18.3}.  The matrix, hidden-continuation
operator, carrier Lie, and signed symplectic bridges are
Proposition~\ref{prop:16.3} and Theorems~\ref{thm:17.1},~\ref{thm:19.1},
and~\ref{thm:20.1}.  The finite proof interface is summarized in
Section~\ref{finite-certificate-protocol}.
\end{proof}

\subsection{The selected point and the second finalist}\label{selected-point-and-second-finalist}

We denote the selected magma by \(\RP\), for \emph{right-row projection}.
The second local finalist is denoted by \(\RC\), for \emph{row-complement}.
The historical support alias \(\mathrm{PAB}\) refers to \(\RP\); the
mathematical text uses \(\RP\) for the selected object.

The two finalists have the same global placement in the \(\mathcal H\)-frontier:
\[
H_{tot}=7020,
\qquad
N_-=0.
\]
They are separated by the directed-edge geometry on
\[
M^\times=\{(r,c):r\ne c\}.
\]
The selected magma realizes the literal continuation/reversal pair
\(\{C,J\}\), while the second finalist realizes the drifted pair
\(\{C^{-1},C^{-1}J\}\).  The final selector is the undrifted finite
\(C/J\) geometry.

\subsection{Information selectors}\label{information-selectors}

The first two selectors are information-theoretic observables on finite
random variables.  The hidden-column access functional is
\[
\mathcal H_{acc}=I(C_{out};C_1,C_2\mid R_1,R_2).
\]
Its zero set is exactly the column-blind cross-row locus: once the ordered
row pair is known, the output column is constant over the hidden input
columns.  This is the Landauer-facing economy selector.

The diagonal selector is the Shannon entropy of the same-point diagonal
output multiset,
\[
H_{diag}=-\sum_s p_s\log_2 p_s.
\]
Its zero set is the constant diagonal sector.  Row-anchor idempotence is
the anchor condition
\[
(r,r)\cdot(r,r)=(r,r)\qquad(r\in S),
\]
and, after diagonal compression, it fixes the constant diagonal to
\(d=000\).  Equivalently, every same-point square then returns to the row
anchor \((r,r)\).  The pure \(C/J\) condition completes the selection by
imposing literal directed-edge head-continuation and reversal.
Appendix~\ref{appendix-information-theoretic-interpretation} records the
information-theoretic reading of these selectors and of the selected magma.

\subsection{Landscape atlases versus selection}\label{landscape-atlases-versus-selection}

The global atlases and the selector have different roles.  The associativity
atlas and the \(\mathcal H\)-frontier atlas map the ambient finite landscape
\(\Omega'\).  The compact selector is a local finite theorem inside that
landscape: after access-zero, diagonal entropy zero, and row-anchor
idempotence, the pure \(C/J\) condition already forces \(\RP\).

A longer explanatory chain is retained because it shows the intermediate
strata:
\[
\Omega'
\xrightarrow{\mathcal H_{acc}=0,
\ H_{diag}=0}
9\times3
\xrightarrow{a\ne b,
\ d=000}
6\times1
\xrightarrow{\min\operatorname{Assoc}_{000}}
2\times1
\xrightarrow{\text{pure }C/J}
1.
\]
This chain displays the \(\RP\)/\(\RC\) ambiguity before the undrifted
\(C/J\) separation.  The compact proof in Theorem~\ref{thm:12.3} is the
shorter logical core.

\subsection{Linearization after selection}\label{linearization-after-selection}

The post-selection part of the paper records the finite passport of
\(\RP\).  The row-owner law says that every nonempty term has output row
equal to the row of its leftmost variable.  The left translations have rank
\(3\) and image a row fiber.  Their positive composition envelope has
\(51\) maps and a three-component row-target product law.

The carrier vector space has a natural matrix-unit identification
\[
k[M]\cong\operatorname{Mat}_3(k),
\qquad e_{(r,c)}\leftrightarrow E_{rc}.
\]
Under this identification the cross-row part of \(\RP\) is represented by
the row-contraction bridge
\[
\mu_{cross}(A,B)=AJ_3B^\top=(A\mathbf1)(B\mathbf1)^\top,
\qquad
E_{ij}J_3E_{kl}^{\top}=E_{ik}.
\]
The carrier-level Lie and signed symplectic structures are developed at
this linearized level.

\subsection{Proof architecture and support}\label{proof-architecture-and-support}

The proof package uses the following finite support modes.

\begin{longtable}{@{}p{0.30\linewidth}p{0.32\linewidth}p{0.30\linewidth}@{}}
\toprule
result family & role & support \\
\midrule
\endhead
normal form, access-zero, diagonal compression, compact \(\RP\) selector & hand finite theorems & text; L2 audit tables record survivor counts \\
associativity range, endpoints, partition function & finite landscape atlas & \texttt{L1/} tables plus L1 verifier \\
global \(\mathcal H\)-frontier & finite frontier theorem & separate \texttt{mathcal\_H/} bundle \\
operator, endomorphism, composition, \(T_2\) structure & post-selection passport & text plus \texttt{L3/} realization and audit tables \\
carrier Lie and signed symplectic structures & bridge layer & \texttt{L3/} bridge audit tables \\
\bottomrule
\end{longtable}

The distributed support corpus keeps the frozen finite layers and the
\(\mathcal H\)-frontier certificate bundle separate:
\[
\texttt{L1/},
\qquad
\texttt{L2/},
\qquad
\texttt{L3/},
\qquad
\texttt{mathcal\_H/}.
\]
The reviewer-facing finite-layer audit is
\begin{verbatim}
tools/run_quick_checks.sh
\end{verbatim}
The \(\mathcal H\)-frontier certificate bundle is kept under
\texttt{mathcal\_H/}; its command-level interface is recorded in
\texttt{SUPPORT\_INDEX.txt}.

\subsection{Positioning}\label{positioning}

The selection layer uses standard finite information observables in the
language of Landauer, Bennett, Shannon, and Cover--Thomas
\cite{Landauer1961,Bennett1973,Bennett1982,Shannon1948,CoverThomas2006}.
The finite magma, term-function, and clone-theoretic context is standard
universal algebra
\cite{Birkhoff1935,Cohn1965,BurrisSankappanavar1981,Post1941,Rosenberg1970,Szendrei1986,Lau2006}.
The same-row rule is the three-point Steiner multiplication, in the
classical nonassociative and quasigroup background
\cite{Bruck1958,DenesKeedwell1974,Pflugfelder1990,ColbournDinitz2007}.
The certificate layer belongs to computer-assisted finite mathematics
\cite{AppelHaken1977,AppelHakenKoch1977,Gonthier2008,McCune1997,Hales2005,HalesEtAl2017}.
The carrier-level matrix, Lie, and symplectic terminology follows standard
conventions
\cite{HornJohnson2012,FultonHarris1991,Hall2015,GuilleminSternberg1990,McDuffSalamon2017}.


\subsection{Notation}\label{notation}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
notation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
meaning
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(S\) & the field \(\mathbb F_3\) \\
\(M\) & the carrier \(S\times S\) \\
\(\sigma\) & the simultaneous shift \((r,c)\mapsto(r+1,c+1)\) \\
\(F_r\) & the row fiber \(\{r\}\times S\) \\
\(M^\times\) & the six off-diagonal points \(\{(r,c):r\ne c\}\) \\
\(\Omega'\) & the landscape satisfying AX0, AX2, and \(\sigma\)-equivariance \\
\((A,B,d)\) & normal coordinates in \(S^{18}\times S^3\) \\
\(h_{a,b}\) & column-blind cross-row rule with anchors \(h(0,1)=a\),
\(h(0,2)=b\) \\
\(\operatorname{Assoc}\) & total number of associative triples in
\(M^3\) \\
\(\operatorname{Assoc}_{000}\) & local associativity statistic after
fixing \(d=000\) \\
\(\mathcal H_{acc}\) & hidden-column access observable \\
\(H_{diag}\) & entropy of the diagonal output multiset \\
\(h_q,H_{tot},N_-\) & hidden-continuation local contribution, total
value, and negative local count \\
\(C,J\) & continuation and reversal maps on \(M^\times\) \\
\(T_2(\RP)\) & binary term-function algebra of \(\RP\) \\
\end{longtable}

\part{The finite landscape}


\section{Carrier, axioms, and normal
form}\label{carrier-axioms-and-normal-form}

\subsection{Carrier and shift}\label{carrier-and-shift}

Let
\[
S=\mathbb F_3,
\qquad
M=S\times S.
\]
All arithmetic in row and column coordinates is in \(S\).  For
\(x=(r,c)\in M\), write
\[
\operatorname{row}(x)=r,
\qquad
\operatorname{col}(x)=c,
\]
and let
\[
F_r=\{r\}\times S
\]
be the row fiber.  We use the cyclic shift
\[
\sigma(r,c)=(r+1,c+1).
\]
Thus \(\sigma\) translates rows and columns simultaneously.

For distinct \(i,j\in S\), write \(\overline{ij}\) for the third element
of \(S\setminus\{i,j\}\).  Equivalently, in \(\mathbb F_3\),
\[
\overline{ij}=-i-j
\qquad (i\ne j).
\]

\subsection{Axioms}\label{axioms}

The landscape \(\Omega'\) consists of binary operations \(\cdot\) on \(M\)
satisfying the following three conditions.

\begin{description}[leftmargin=3.2em,style=nextline]
\item[AX0, left-row output.]
For all \(x,y\in M\),
\[
\operatorname{row}(x\cdot y)=\operatorname{row}(x).
\]
Equivalently, \(x\cdot y\in F_{\operatorname{row}(x)}\).

\item[\(\sigma\)-equivariance.]
For all \(x,y\in M\),
\[
\sigma(x\cdot y)=\sigma(x)\cdot \sigma(y).
\]

\item[AX2, same-row off-diagonal Steiner rule.]
For all \(r,i,j\in S\) with \(i\ne j\),
\[
(r,r+i)\cdot(r,r+j)=(r,r+\overline{ij}).
\]
Equivalently, in absolute column notation,
\[
(r,c_1)\cdot(r,c_2)=(r,\overline{c_1c_2})
\qquad(c_1\ne c_2).
\]
This is the Steiner multiplication on a three-point row fiber
\cite{Pflugfelder1990,ColbournDinitz2007}.
\end{description}

AX0 fixes only the output row.  The output column remains free in the
cross-row branch and on the same-point diagonal branch.

\subsection{Normal coordinates}\label{normal-coordinates}

It is convenient to use row-relative coordinates.  A point of row \(r\) is
written as \((r,r+i)\), where \(i\in S\) is the column offset from the row.

A normal coordinate is a triple
\[
(A,B,d)\in S^9\times S^9\times S^3,
\]
where
\[
A=(A_{ij})_{i,j\in S},
\qquad
B=(B_{ij})_{i,j\in S},
\qquad
 d=(d_0,d_1,d_2).
\]
The associated multiplication is defined by the four formulas
\[
(r,r+i)\cdot(r+1,r+1+j)=(r,r+A_{ij}),
\]
\[
(r,r+i)\cdot(r+2,r+2+j)=(r,r+B_{ij}),
\]
\[
(r,r+i)\cdot(r,r+i)=(r,r+d_i),
\]
and, for \(i\ne j\),
\[
(r,r+i)\cdot(r,r+j)=(r,r+\overline{ij}).
\]
The first two arrays record the two ordered cross-row representatives
\((0,1)\) and \((0,2)\).  The vector \(d\) records the same-point diagonal.
The fourth formula is AX2.

Conversely, any operation satisfying AX0, \(\sigma\)-equivariance, and AX2
has a unique coordinate triple \((A,B,d)\): after applying a power of
\(\sigma\), every cross-row product reduces to one of the two row-pair
representatives above, and every same-point product reduces to one of the
three diagonal coordinates.

\subsection{Normal form theorem}\label{normal-form-theorem}

\begin{theorem}
\label{thm:1.1}
The AX0+AX2, \(\sigma\)-equivariant landscape has normal form
\[
\Omega'\cong S^{18}\times S^3\cong S^{21}.
\]
Consequently
\[
|\Omega'|=3^{21}=10\,460\,353\,203.
\]
\end{theorem}

\begin{proof}
AX0 fixes the output row of every product.  AX2 fixes all same-row
off-diagonal output columns.  The same-point diagonal branch has three
row-relative coordinates \(d_0,d_1,d_2\).  For cross-row products,
\(\sigma\)-equivariance reduces the ordered row-pair data to the two
representatives \((0,1)\) and \((0,2)\).  Each representative has
\(3\times3\) column-offset inputs, hence contributes nine independent
coordinates.  Thus the cross-row branch contributes \(18\) coordinates and
the diagonal branch contributes \(3\) coordinates.

The coordinate construction above gives a product satisfying AX0,
\(\sigma\)-equivariance, and AX2 for every \((A,B,d)\in S^9\times S^9
\times S^3\), and the preceding paragraph recovers these coordinates
uniquely from any such product.  Hence \(\Omega'\cong S^{21}\), and its
cardinality is \(3^{21}\).
\end{proof}

\section{Effective symmetry and orbit
count}\label{effective-symmetry-and-orbit-count}

The oriented system used in the selection theorem fixes the generator
\(s\mapsto s+1\) through
\(\sigma(r,c)=(r+1,c+1)\).  For landscape compression, one may also use the
normalizer of this cyclic shift in \(\operatorname{Sym}(S)\).  Translations
\(s\mapsto s+\beta\) commute with the shift and act trivially on the
row-relative normal coordinates.  The remaining normalizer class is
represented by the reflection \(\nu(s)=-s\), which conjugates
\(\sigma\) to \(\sigma^{-1}\).  Since \(\sigma\)-equivariance is equivalent
to \(\sigma^{-1}\)-equivariance, this reflection preserves \(\Omega'\) as a
landscape-compression symmetry.

On normal coordinates it induces the involution
\[
\iota(A,B,d)=(A',B',d')
\]
with indices in \(S\) and
\[
A'_{ij}=-B_{-i,-j},
\qquad
B'_{ij}=-A_{-i,-j},
\qquad
d'_i=-d_{-i}.
\]
We denote the resulting residual effective action by
\[
G_{\mathrm{eff}}=\langle\iota\rangle\cong C_2.
\]
On the diagonal coordinate sector \(\Delta=S^3\), the same residual
reflection gives the compression
\[
|\Delta/G_{\mathrm{eff}}|=\frac{3^3+3}{2}=15.
\]

\begin{proposition}
\label{prop:2.1}
The effective orbit count of the full
normal-coordinate landscape is
\[
|\Omega'/G_{\mathrm{eff}}|=\frac{3^{21}+3^{10}}2.
\]
\end{proposition}

\begin{proof}
Burnside's lemma applies to the displayed involution.  The identity
contributes \(3^{21}\).  For a fixed point of \(\iota\), the nine entries
of \(A\) are free and then determine \(B\) by
\(B_{ij}=-A_{-i,-j}\).  The diagonal equations are
\[
d_0=-d_0,
\qquad
d_1=-d_2,
\qquad
d_2=-d_1,
\]
so \(d_0=0\) and \(d_1\) is free.  Hence the fixed locus has
\(9+1=10\) ternary degrees of freedom and size \(3^{10}\).  Therefore
\[
|\Omega'/G_{\mathrm{eff}}|=
\frac{3^{21}+3^{10}}2.
\]
\end{proof}

The orbit count is a landscape compression.  The residual reflection is
not used to identify the two directed-edge absorption geometries in the
selection theorem; the pure \(C/J\) maps are defined intrinsically on the
directed-edge set in Section~\ref{pure-cj-and-the-selection-theorem}.

\section{Hidden-column access}\label{hidden-column-access}

The first selection functional is information-theoretic. It measures
whether the cross-row rule has functional dependence on hidden column
coordinates after the row pair is known. Its vanishing is the
Landauer-facing condition that the cross-row channel admits a
realization through the row data alone.

\subsection{Column-blind cross-row
rules}\label{column-blind-cross-row-rules}

A cross-row output rule is \textbf{column-blind} if its output column
depends only on the two row coordinates: \[
g(r_1,c_1,r_2,c_2)=h(r_1,r_2).
\] Equivalently, after the ordered row pair is fixed, the output column
is constant over the nine column inputs.

By \(\sigma\)-equivariance, a column-blind rule is determined by two
values \[
a=h(0,1),
\qquad
b=h(0,2).
\] We call \((a,b)\) the cross-row anchors and write \(h_{a,b}\) for the
corresponding rule. The remaining ordered row pairs are forced by
\(\sigma\)-shift: \[
h(r,r+1)=r+a,
\qquad
h(r,r+2)=r+b
\] with arithmetic in \(\mathbb F_3\). There are \(3^2=9\) column-blind
cross-row rules.

\subsection{Hidden-column access as a Landauer-facing
observable}\label{hidden-column-access-as-a-landauer-facing-observable}

Let \[
\Sigma_{cross}=\{(r_1,c_1,r_2,c_2)\in S^4:r_1\ne r_2\}
\] with the uniform distribution, and let \[
C_{out}=g(R_1,C_1,R_2,C_2)
\] be the output-column random variable. Define the hidden-column access
observable by \[
\boxed{
\mathcal H_{acc}(g)=I(C_{out};C_1,C_2\mid R_1,R_2).
}
\] Logarithms are base \(2\), so the value is measured in bits.

This is a structural information observable. It measures how much
hidden-column information is functionally retained in the cross-row
output once the row pair is fixed. The Landauer reading is that hidden
column information which is read and later discarded by an irreversible
implementation carries thermodynamic cost
\cite{Landauer1961,Bennett1973,Bennett1982}. Therefore
\(\mathcal H_{acc}=0\) selects exactly those cross-row rules for which
that cost is structurally avoidable: the output is computed through
the public row interface.

The zero condition is used in this implementation-independent structural
sense: the column-read burden can be avoided altogether.

\subsection{Access-zero theorem}\label{access-zero-theorem}

\begin{theorem}
\label{thm:3.1}
For a cross-row rule \(g\), \[
\mathcal H_{acc}(g)=0
\quad\Longleftrightarrow\quad
 g(r_1,c_1,r_2,c_2)=h(r_1,r_2).
\] There are exactly nine column-blind cross-row rules.
\end{theorem}

\begin{proof}
Conditional mutual information vanishes exactly when \[
C_{out}\perp (C_1,C_2)\mid (R_1,R_2).
\] Since \(C_{out}\) is a deterministic function of
\((R_1,C_1,R_2,C_2)\), conditional independence is equivalent to
constancy in \((C_1,C_2)\) for every fixed ordered row pair
\((R_1,R_2)\). That is precisely column-blindness.

The two representative cross-row row pairs \((0,1)\) and \((0,2)\) each
receive one output-column value in \(S\), giving \(3^2=9\)
\(\sigma\)-equivariant column-blind rules.
\end{proof}

This theorem is a landscape result and also the first reduction in the
selection chain.

\section{Associativity master
formula}\label{associativity-master-formula}

\subsection{The associativity
statistic}\label{the-associativity-statistic}

For \(x\in\Omega'\), define \[
\operatorname{Assoc}(x)=
\#\{(u,v,w)\in M^3:(u\cdot v)\cdot w=u\cdot(v\cdot w)\}.
\] Since \(|M|=9\), this statistic lies between \(0\) and \(729\).

\begin{theorem}
\label{thm:4.1}
In normal coordinates, \[
\operatorname{Assoc}=3\,rawAssoc,
\] where \(rawAssoc\) is a normalized sum over \(3^5=243\) triples and
decomposes as \[
rawAssoc=rawRRR+rawRRS+rawRSR+rawSRR+rawDIST.
\] The five blocks correspond to the row patterns \[
RRR,
\qquad RRS,
\qquad RSR,
\qquad SRR,
\qquad DIST.
\]
\end{theorem}

\begin{proof}
By \(\sigma\)-equivariance the first row coordinate of a
triple may be normalized to \(0\). The two remaining row variables and
the three column variables give \(3^5\) normalized cases; the factor
\(3\) restores the three translates. The displayed blocks are the five
equality patterns of the three rows.
\end{proof}

\subsection{The same-row block}\label{the-same-row-block}

The all-same-row block depends only on the diagonal vector \(d\). Let \[
E(d)=\#\{(i,j):i\ne j,
\ d_i=d_j\},
\qquad
N(d)=\#\{(i,j):i\ne j,
\ d_i=j,
\ d_j=j\}.
\]

\begin{lemma}
\label{lem:4.2}
The same-row contribution is \[
rawRRR(d)=9+E(d)+2N(d).
\] The four diagonal classes give raw same-row values \[
9,\qquad 11,\qquad 13,\qquad 19.
\] In particular, \[
rawRRR(000)=19.
\]
\end{lemma}

\begin{proof}
In one row, unequal column pairs use the Steiner rule
and equal pairs use \(d\). The \(27\) triples split into a baseline
family plus the two diagonal-sensitive families counted by \(E(d)\) and
\(N(d)\). This gives the formula and the four class values.
\end{proof}

\subsection{\texorpdfstring{Column-blind profile at
\(d=000\)}{Column-blind profile at d=000}}\label{column-blind-profile-at-d000}

Let \(h_{a,b}\) be the column-blind rule \[
h(r,r+1)=r+a,
\qquad
h(r,r+2)=r+b.
\] At \(d=000\), same-point products satisfy \((r,c)^2=(r,r)\).

\begin{proposition}
\label{prop:4.3}
On the column-blind, \(d=000\) stratum, \[
\begin{aligned}
&(rawRRR,rawRRS,rawRSR,rawSRR,rawDIST)(a,b,000)\\
&\quad =
\bigl(19,\, 9[a=0]+9[b=0],\, 9[a=0]+9[b=0],\\
&\hspace{7em}54,\,54[a=b]\bigr).
\end{aligned}
\] Consequently \[
\boxed{
\operatorname{Assoc}_{000}(h_{a,b})
=
3\left(73+18[a=0]+18[b=0]+54[a=b]\right).
}
\]
\end{proposition}

\begin{proof}
The \(RRR\) value is Lemma 4.2. In the \(RRS\) and
\(RSR\) patterns, the same-row step must return the cross-row anchor; at
\(d=000\) this contributes \(9\) exactly when the relevant anchor is
\(0\). In \(SRR\), the final cross-row output ignores same-row column
processing, giving all \(54\) normalized triples. In \(DIST\), the two
association orders compare the two cross-row anchors, so all \(54\)
triples survive exactly when \(a=b\).
\end{proof}

The nine column-blind rules split as follows:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2400}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3000}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2600}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
representatives
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
profile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\operatorname{Assoc}_{000}\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
anchored trivial & \((0,0)\) & \((19,18,18,54,54)\) & \(489\) \\
off-anchor trivial & \((1,1),(2,2)\) & \((19,0,0,54,54)\) & \(381\) \\
partly anchored nondegenerate & \((0,1),(0,2),(1,0),(2,0)\) &
\((19,9,9,54,0)\) & \(273\) \\
fully off-anchor nondegenerate & \((1,2),(2,1)\) & \((19,0,0,54,0)\) &
\(219\) \\
\end{longtable}

The fourth class consists of \(\RP\) and \(\RC\).

\section{Global associativity range}\label{global-associativity-range}

\subsection{Exact range}\label{exact-range}

\begin{lemma}[Associativity atlas certificate protocol]
\label{lem:5.1}
The
global associativity atlas is used through five finite obligations.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{A1, block reduction.} The associativity statistic is reduced
  by \(\sigma\)-normalization to the five row-pattern blocks \[
  rawAssoc=rawRRR+rawRRS+rawRSR+rawSRR+rawDIST
  \] of Theorem 4.1.
\item
  \textbf{A2, diagonal decomposition.} The same-row block is evaluated
  from the diagonal vector by Lemma 4.2, giving the four possible values
  \[
  9,\quad 11,\quad 13,\quad 19.
  \]
\item
  \textbf{A3, fixed-diagonal cross-range certificates.} For each
  diagonal class, the certificate verifies the exact range of \[
  rawCross=rawRRS+rawRSR+rawSRR+rawDIST.
  \]
\item
  \textbf{A4, endpoint witnesses and endpoint counts.} The endpoint
  certificates verify the normal-coordinate witnesses, endpoint counts,
  and orbit representatives for \(\operatorname{Assoc}=63\) and
  \(\operatorname{Assoc}=597\).
\item
  \textbf{A5, distribution consistency.} The partition-function
  certificate gives the full coefficient table for \[
  Z(q)=\sum_{x\in\Omega'}q^{\operatorname{Assoc}(x)}
  \] and checks total mass, range, moments, and mode.
\end{enumerate}

Any certificate satisfying A1--A5 proves the global associativity range,
endpoint loci, and partition-function claims of Sections 5--6. This is a
certified atlas statement for the ambient landscape; the compact \(\RP\)
uniqueness proof in Theorem 12.3 is the separate hand spine in Part II.
\hfill$\square$
\end{lemma}

\begin{theorem}[Global associativity atlas]
\label{thm:5.2}
On \(\Omega'\), \[
21\le rawAssoc\le199.
\] Equivalently, \[
63\le\operatorname{Assoc}\le597.
\]

The finite certificate is organized by diagonal class:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2000}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2667}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
diagonal class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
rawRRR
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
certified rawCross range
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
rawAssoc consequence
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
permutation & 9 & 12..180 & 21..189 \\
two-valued, no fixed point & 11 & 10..162 & 21..173 \\
two-valued, with fixed point & 13 & 12..174 & 25..187 \\
constant & 19 & 12..180 & 31..199 \\
\end{longtable}

Here \[
rawCross=rawRRS+rawRSR+rawSRR+rawDIST.
\] The endpoint values are therefore \(rawAssoc=21\) and
\(rawAssoc=199\), or \(\operatorname{Assoc}=63\) and
\(\operatorname{Assoc}=597\).
\end{theorem}

\begin{proof}
The hand-readable reduction is Theorem 4.1 plus Lemma
4.2. After that reduction, the global claim is exactly the A3
cross-range obligation of Lemma 5.1. Adding the certified cross-row
ranges to the four possible same-row values gives the displayed global
range.

The endpoint witnesses, endpoint counts, orbit representatives, and
partition-function consistency are the A4--A5 obligations of Lemma 5.1.
They are recorded in the \texttt{L1/tables/} atlas files listed in
\texttt{SUPPORT\_INDEX.txt} and checked by
\texttt{L1/scripts/verify\_layer1\_v3\_final.py}.
\end{proof}

\subsection{Global and local associativity
roles}\label{global-and-local-associativity-roles}

\(\RP\) has \[
\operatorname{Assoc}(\RP)=219,
\] while the landscape minimum is \(63\). The global statistic maps the
ambient landscape. The selection argument later uses the conditional
local statistic \(\operatorname{Assoc}_{000}\), after information
compression and anchoring, to expose the final two-candidate ambiguity.

\section{Endpoint loci and partition
function}\label{endpoint-loci-and-partition-function}

\subsection{Endpoint loci}\label{endpoint-loci}

\begin{theorem}
\label{thm:6.1}
The associativity endpoint loci have sizes \[
|\mathcal M_{63}|=24,
\qquad
|\mathcal M_{63}/\operatorname{Sym}(S)|=12,
\] and \[
|\mathcal M_{597}|=6,
\qquad
|\mathcal M_{597}/\operatorname{Sym}(S)|=3.
\] The minimum endpoint block profiles are \[
(11,2,2,2,4)
\quad\text{and}\quad
(9,2,2,4,4),
\] while the maximum endpoint block profile is \[
(19,36,36,54,54).
\]
\end{theorem}

\begin{proof}
This is the A4 endpoint part of Lemma 5.1. The endpoint
certificates record the exact solution counts by diagonal sector, orbit
representatives, and explicit witnesses in normal coordinates. The
minimum occurs in the permutation and two-valued/no-fixed diagonal
classes. The maximum occurs in the constant diagonal class. The block
profiles are obtained by evaluating the five normalized associativity
blocks on the endpoint witnesses.
\end{proof}

\subsection{Exact associativity partition
function}\label{exact-associativity-partition-function}

\begin{theorem}
\label{thm:6.2}
The exact polynomial \[
Z(q)=\sum_{x\in\Omega'}q^{\operatorname{Assoc}(x)}
\] has \(167\) nonzero coefficients and total mass \[
Z(1)=3^{21}.
\] Its first two moments are \[
\mathbb E[\operatorname{Assoc}]=245,
\qquad
\operatorname{Var}(\operatorname{Assoc})=\frac{27820}{27}.
\] The mode occurs at \[
\operatorname{Assoc}=237
\] with coefficient \[
426,317,976.
\]
\end{theorem}

\begin{proof}
This is the A5 partition-function part of Lemma 5.1. The
coefficient certificate in \texttt{L1/tables/} is the exact finite
distribution of \(\operatorname{Assoc}\) on \(\Omega'\).  The total mass,
range, mean, variance, and mode are computed directly from that table and
checked by \texttt{L1/scripts/verify\_layer1\_v3\_final.py}.
\end{proof}

\section{\texorpdfstring{Hidden continuation and the \(\mathcal H\)-frontier}{Hidden continuation and the H-frontier}}\label{hidden-continuation}

\subsection{Local distance reduction}\label{local-distance-reduction}

The hidden-continuation observable is a finite path-information defect.
The continuation side keeps the two-step route data, while the
left-translation side compares only the collapsed endpoint operators.
Thus \(h_q/2\) is the signed Hamming contrast between a path-resolved
comparison and an endpoint-resolved comparison. Negative local values
are precisely the local cancellation failures counted by \(N_-\). The
normalization \(H_{tot}=6rawH\) is the \(\sigma\)-orbit normalization
used by the \(\mathcal H\)-frontier certificate.

Its local index is
\[
q=(b,t,a,e,f)\in S^5.
\]
For a completed normal-form point \(x\in\Omega'\), let \(M_s\) denote
the local table read: \(M_1\) and \(M_2\) are the two cross-row tables,
and \(M_0\) is the same-row diagonal/complement rule. All row and column
arithmetic below is in \(S=\mathbb F_3\). Define
\[
xy=M_b(a,e),
\qquad
yz=M_{t-b}(e-b,f-b),
\qquad
z_R=b+yz,
\]
and
\[
ep_L=M_t(xy,f),
\qquad
ep_R=M_b(a,z_R).
\]

For \(c\in S\), define the collapsed left-translation signature
\[
L_c(i,j)=M_i(c,j)
\qquad ((i,j)\in S^2).
\]
For \(s,\alpha,z\in S\), define the stepwise continuation signature
\[
C_{s,\alpha,z}(i,j)
=
M_s\!\left(
\alpha,\,
s+M_{i-s}(z-s,j-s)
\right)
\qquad ((i,j)\in S^2).
\]
The associated Hamming distances are
\[
D_L(c,c')
=
\#\{(i,j)\in S^2:L_c(i,j)\ne L_{c'}(i,j)\},
\]
and
\[
D_C(s,\alpha,z;s',\alpha',z')
=
\#\{(i,j)\in S^2:
C_{s,\alpha,z}(i,j)\ne C_{s',\alpha',z'}(i,j)\}.
\]

\paragraph{Direct local definition.}
For each of the nine test pairs \((i,j)\in S^2\), set
\[
\delta^C_{ij}(q;x)
=
\mathbf 1\!\left\{
C_{t,xy,f}(i,j)\ne C_{b,a,z_R}(i,j)
\right\},
\]
and
\[
\delta^L_{ij}(q;x)
=
\mathbf 1\!\left\{
L_{ep_L}(i,j)\ne L_{ep_R}(i,j)
\right\}.
\]
The nine direct local summands are
\[
\Delta_{ij}(q;x)=2\bigl(\delta^C_{ij}(q;x)-\delta^L_{ij}(q;x)\bigr),
\]
displayed as
\[
\begin{array}{c|ccc}
 & j=0 & j=1 & j=2\\
\hline
i=0 & \Delta_{00} & \Delta_{01} & \Delta_{02}\\
i=1 & \Delta_{10} & \Delta_{11} & \Delta_{12}\\
i=2 & \Delta_{20} & \Delta_{21} & \Delta_{22}
\end{array}
\]
and the direct hidden-continuation contribution is
\[
h_q(x)
=
\sum_{(i,j)\in S^2}\Delta_{ij}(q;x).
\]

\begin{theorem}[Local distance reduction]
\label{thm:7.1}
For every \(x\in\Omega'\) and \(q\in S^5\),
\[
\boxed{
h_q(x)=
2\left(
D_C(t,xy,f;b,a,z_R)-D_L(ep_L,ep_R)
\right).
}
\]
Consequently the local direct definition and the Hamming-distance formula
define the same global observables \(rawH\) and \(H_{tot}\).
\end{theorem}

\begin{proof}
By definition, \(D_C(t,xy,f;b,a,z_R)\) is the sum of the nine indicators
\(\delta^C_{ij}(q;x)\), and \(D_L(ep_L,ep_R)\) is the sum of the nine
indicators \(\delta^L_{ij}(q;x)\). Subtracting these two Hamming
distances and multiplying by \(2\) gives exactly the displayed sum of
the nine direct local summands \(\Delta_{ij}\). The symbolic reduction is recorded by the \(\mathcal H\)-frontier
certificate handles \texttt{HF-RED} and \texttt{HF-CERT-IF}.
\end{proof}

Then
\[
rawH(x)=\sum_{q\in S^5}\frac{h_q(x)}2,
\qquad
H_{tot}(x)=6\,rawH(x)=3\sum_{q\in S^5}h_q(x),
\]
and
\[
\boxed{
N_-(x)=3\,\#\{q\in S^5:h_q(x)<0\}.
}
\]
Thus purity is the finite system
\[
N_-(x)=0
\quad\Longleftrightarrow\quad
h_q(x)\ge0\quad(q\in S^5).
\]

\subsection{Global frontier theorem}\label{global-frontier-theorem}

The global \(\mathcal H\)-frontier theorem is supported by four finite
certificate obligations.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
handle
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
obligation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
result
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
H1 & unrestricted full-landscape range & \(rawH\in[-378,1217]\),
endpoint counts \(8,12\) \\
H2 & pure-frontier exclusion &
\((h_q\ge0\ \forall q)\wedge rawH\ge1171\) is UNSAT \\
H3 & pure-frontier witnesses & \(\RP\) and \(\RC\) have
\(rawH=1170\), \(N_-=0\) \\
H4 & signed-cancellation spike & exactly twelve points above the pure
frontier, all with \((rawH,N_-)=(1217,3)\) \\
\end{longtable}

\begin{theorem}[Global frontier of \(\mathcal H\)]
\label{thm:7.2}
On
\(\Omega'=S^{21}\):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The unrestricted range is \[
  \boxed{
  \min H_{tot}=-2268,
  \qquad
  \max H_{tot}=7302.
  }
  \] Equivalently, \(rawH\in[-378,1217]\). The endpoint multiplicities
  are \(8\) and \(12\).
\item
  The pure frontier is \[
  \boxed{
  \max\{H_{tot}:N_-=0\}=7020.
  }
  \]
\item
  The region above the pure frontier is exactly the unrestricted maximum
  layer: \[
  \boxed{
  \{H_{tot}>7020\}=\{H_{tot}=7302\}.
  }
  \] This layer has \(12\) points, all with \[
  rawH=1217,
  \qquad H_{tot}=7302,
  \qquad N_-=3,
  \] \[
  H_{pos}=7308,
  \qquad H_{\mathrm{neg\_abs}}=6,
  \qquad h_{\mathrm{loc,min}}=-2,
  \qquad h_{\mathrm{loc,max}}=18.
  \]
\end{enumerate}
\end{theorem}

\begin{proof}
The analytic reduction is Theorem 7.1. H1 covers all
\(3^{21}\) normal-form words and gives the unrestricted range. For the
pure frontier, \(N_-=0\) is the system \(h_q\ge0\) for all \(q\in S^5\),
and \(H_{tot}>7020\) is equivalent to \(rawH\ge1171\). H2 proves \[
(h_q\ge0\ \forall q\in S^5)\wedge rawH\ge1171
\] UNSAT on all \(18540\) depth-9 live nodes: \[
18540/18540\ \mathrm{UNSAT},
\qquad SAT=0,
\qquad UNKNOWN=0.
\] H3 supplies pure witnesses at \(rawH=1170\), and H4 classifies the
full layer above the pure threshold as the twelve impure maximizers.
\end{proof}

\subsection{\texorpdfstring{\(\RP\) and \(\RC\) on the pure frontier}{RP and RC on the pure frontier}}\label{rp-and-rc-on-the-pure-frontier}

\begin{corollary}
\label{cor:7.3}
\(\RP\) and \(\RC\) satisfy \[
H_{tot}=7020,
\qquad rawH=1170,
\qquad N_-=0.
\] Thus both lie on the global pure \(\mathcal H\)-frontier.
\end{corollary}

\begin{proof}
Direct evaluation of \[
\RP=111111111222222222000,
\qquad
\RC=222222222111111111000
\] gives the displayed values, and Theorem 7.2 proves that no pure point
has larger \(H_{tot}\).
\end{proof}

Part I has now supplied the ambient finite map: normal form, global
associativity, and the \(\mathcal H\)-frontier. Part II turns from
atlas data to the intrinsic selector. Its proof uses the access-zero and
diagonal-compression reductions, then the directed-edge absorption
geometry; the frontier theorem supplies the global placement of the two
selection finalists.

\part{Intrinsic finite selection}

\section{Selection core}\label{selection-core}

The selection theorem is the finite hand spine of the paper.  It uses four
internal conditions:
\[
\boxed{
\mathcal H_{acc}=0,
\qquad
H_{diag}=0,
\qquad
\text{row-anchor idempotence},
\qquad
\text{pure }C/J.
}
\]
Their order follows the row/column geometry of \(\Omega'\).  The first
condition makes the cross-row branch column-blind, so it has anchors
\((a,b)\in S^2\).  The second condition compresses the same-point diagonal
to a constant triple, and row-anchor idempotence fixes this triple to
\(d=000\).  At that point the absorption geometry on the six directed edges
\(M^\times\) reads the remaining cross-row anchors directly; the literal
continuation/reversal criterion forces \((a,b)=(1,2)\).  Thus the compact
core selects the right-row projection magma \(\RP\).

In normal coordinates, the selected point and the second diagnostic finalist
are
\[
\RP=(a,b,d)=(1,2,000),
\qquad
\RC=(a,b,d)=(2,1,000).
\]
The distinction between them is geometric: \(\RP\) realizes the undrifted
pair \(\{C,J\}\) on directed edges, whereas \(\RC\) realizes
\(\{C^{-1},C^{-1}J\}\).

The longer explanatory chain is retained as a diagnostic path through the
same finite landscape:
\[
\Omega'
\xrightarrow{\mathcal H_{acc}=0,
\ H_{diag}=0}
9\times3
\xrightarrow{a\ne b,
\ d=000}
6\times1
\xrightarrow{\min\operatorname{Assoc}_{000}}
2\times1
\xrightarrow{\text{pure }C/J}
1.
\]
It records how local path-dependence exposes the \(\RP\)/\(\RC\) ambiguity
before the undrifted \(C/J\) criterion separates the two finalists.  The
logical core remains Theorem~\ref{thm:12.3}.

\section{Diagonal compression}\label{diagonal-compression}

The access-zero theorem has already compressed the cross-row sector to
the nine column-blind rules \(h_{a,b}\). The remaining
information-compression step acts on the same-point diagonal data.

For a diagonal triple \[
d=(d_0,d_1,d_2)\in S^3,
\] let \[
p_s=\frac13\#\{i:d_i=s\}.
\] Define the diagonal Shannon observable by \[
\boxed{
H_{diag}(d)=-\sum_{s\in S}p_s\log_2 p_s.
}
\] The possible levels are:

\begin{longtable}[]{@{}lrr@{}}
\toprule\noalign{}
diagonal type & multiplicities & \(H_{diag}\) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
constant & \((3,0,0)\) & \(0\) \\
two-valued & \((2,1,0)\) & \(\log_2 3-\frac23\) \\
permutation & \((1,1,1)\) & \(\log_2 3\) \\
\end{longtable}

\begin{theorem}
\label{thm:9.1}
For diagonal triples, \[
H_{diag}(d)=0
\quad\Longleftrightarrow\quad
 d_0=d_1=d_2.
\] Thus diagonal compression reduces the diagonal sector from \(27\)
choices to \[
000,
\qquad 111,
\qquad 222.
\]
\end{theorem}

\begin{proof}
Entropy of a finite distribution vanishes exactly when
the distribution is concentrated on one symbol. For the empirical
distribution of \((d_0,d_1,d_2)\), this is precisely constancy of the
diagonal triple.
\end{proof}

Together with the access-zero theorem of Section 3, this reduces the
selection domain to \(9\times3\).

\section{Diagnostic anchors and local
path-dependence}\label{anchors-and-local-path-dependence}

This section records the explanatory branch of the selection story.  The
condition \(a\ne b\) and the local statistic \(\operatorname{Assoc}_{000}\)
are useful diagnostics for the finite landscape: they expose the two
nearby finalists before the directed-edge criterion makes the final choice.

\subsection{Nondegenerate diagnostic anchors}\label{nondegenerate-anchors}

\begin{proposition}
\label{prop:10.1}
Among column-blind cross-row rules
\(h_{a,b}\), the condition \[
a\ne b
\] leaves exactly six rules.
\end{proposition}

\begin{proof}
There are nine pairs \((a,b)\in S^2\). The three
diagonal pairs \((a,a)\) identify the two cross-row representatives under a
single anchor. The remaining \(3^2-3=6\) pairs are the nondegenerate rules.
\end{proof}

\begin{proposition}
\label{prop:10.2}
After diagonal compression, row-anchor
idempotence selects the unique constant diagonal \[
d=000.
\]
\end{proposition}

\begin{proof}
The compressed diagonal sector is \(\{000,111,222\}\).
Row-anchor idempotence is the condition
\[
(r,r)\cdot(r,r)=(r,r)\qquad(r\in S).
\]
On the compressed diagonal sector \(d=sss\), this condition gives \(s=0\).
Thus the unique compressed diagonal compatible with row-anchor idempotence
is \(d=000\).
\end{proof}

\subsection{\texorpdfstring{Local associativity at
\(d=000\)}{Local associativity at d=000}}\label{local-associativity-at-d000}

\begin{proposition}
\label{prop:10.3}
With \(d=000\) fixed, the six nontrivial
column-blind rules have local associativity values \[
\operatorname{Assoc}_{000}=219
\] for exactly two rules, and \[
\operatorname{Assoc}_{000}=273
\] for the remaining four. The two local minima are \(\RP\) and
\(\RC\).

The local-minimum profile is \[
(19,0,0,54,0),
\] whereas the four discarded rules have profile \[
(19,9,9,54,0).
\] The difference comes from the additional \(RRS\) and \(RSR\)
coincidences in the discarded rules.
\end{proposition}

\begin{proof}
Proposition 4.3 gives \[
\operatorname{Assoc}_{000}(h_{a,b})
=
3\left(73+18[a=0]+18[b=0]+54[a=b]\right).
\] On the nondegenerate stratum \(a\ne b\), this value is minimized
exactly when neither anchor is \(0\), i.e.~at \((a,b)=(1,2)\) and
\((2,1)\). These are \(\RP\) and \(\RC\), and both have value
\(219\). The four remaining nondegenerate pairs contain one zero anchor
and have value \(273\). The corresponding \texttt{L2/tables/} path-dependence table and
\texttt{L2/scripts/verify\_layer2\_selection\_final.py} record the same
six-row audit.
\end{proof}

This step belongs to the diagnostic chain.  It shows why the two
rules \(\RP\) and \(\RC\) are the natural local finalists after information
compression and diagonal anchoring.

\section{Directed-edge geometry}\label{directed-edge-geometry}

Let \[
M^\times=\{(r,c)\in M:r\ne c\}
\] be the six-point directed-edge set. Each \((r,c)\in M^\times\) is
interpreted as the directed edge from \(r\) to \(c\) in the triangle on
vertex set \(S\).

Define \[
C(r,c)=(c,\overline{rc}),
\qquad
J(r,c)=(c,r).
\] The map \(C\) advances a directed edge by making the old head the new tail and passing through the third vertex, while
\(J\) reverses the directed edge.

\begin{theorem}
\label{thm:11.1}
On \(M^\times\), \[
C^3=1,
\qquad
J^2=1,
\qquad
JCJ=C^{-1}.
\] Thus \(C\) and \(J\) generate the dihedral geometry of the directed
triangle.
\end{theorem}

\begin{proof}
Applying \(C\) three times traverses \[
(r,c)\mapsto(c,\overline{rc})\mapsto(\overline{rc},r)\mapsto(r,c).
\] Applying \(J\) twice reverses the directed edge twice. Finally, \[
JCJ(r,c)=JC(c,r)=J(r,\overline{rc})=(\overline{rc},r)=C^{-1}(r,c).
\] The \texttt{L3/} directed-edge audit tables record these finite relation
checks.
\end{proof}

\section{\texorpdfstring{Pure \(C/J\) and the selection
theorem}{Pure C/J and the selection theorem}}\label{pure-cj-and-the-selection-theorem}

\subsection{Intrinsicity of the pure \texorpdfstring{\(C/J\)}{C/J} criterion}\label{intrinsicity-of-the-pure-cj-criterion}

The six-point set \(M^\times\) is canonically the directed-edge set of the
three-vertex row alphabet.  For an edge \(x=(r,c)\), call \(r\) its tail,
\(c\) its head, and \(t=\overline{rc}\) the third vertex.  Then
\[
C(r,c)=(c,t),
\qquad
J(r,c)=(c,r).
\]
Thus
\[
\{C(x),J(x)\}
=
\{y\in M^\times:\operatorname{tail}(y)=\operatorname{head}(x)\}.
\]
The pure \(C/J\) condition is therefore the head-row absorption rule:
for every directed edge \(x\), the absorbed neighbours are exactly the two
directed edges whose tail is the head of \(x\).

This is not a choice of an external cyclic orientation.  For every relabelling
\(\rho\in\operatorname{Sym}(S)\), acting on directed edges by
\(\rho_\times(r,c)=(\rho(r),\rho(c))\), one has
\[
\rho_\times C=C\rho_\times,
\qquad
\rho_\times J=J\rho_\times.
\]
Indeed, \(\rho\) sends the third vertex of \(\{r,c\}\) to the third vertex
of \(\{\rho(r),\rho(c)\}\).  Hence the undrifted pair \(\{C,J\}\) is an
intrinsic directed-edge structure.  The drifted pair
\(\{C^{-1},C^{-1}J\}\) is genuinely different: its two outputs have tail
\(\overline{rc}\), the third vertex, rather than tail \(c\), the head of
\((r,c)\).

\subsection{Absorption and absorption
transitions}\label{absorption-and-absorption-transitions}

For \(x\in M^\times\), define the absorption set \[
\operatorname{Abs}(x)=\{y\in M^\times\setminus\{x\}:x\cdot y=x\}.
\] For a column-blind rule \(h_{a,b}\) at \(d=000\), same-row right
arguments never absorb an off-diagonal \(x\). Thus absorption is
governed by the cross-row equation \[
h(r_x,r_y)=c_x.
\]

A candidate has a two-transition absorption geometry if every
\(x\in M^\times\) has exactly two absorbed neighbours. In that case the
two absorbed-neighbour maps define \(\operatorname{AbsTrans}\). The pure
\(C/J\) criterion requires \[
\operatorname{AbsTrans}=\{C,J\}
\] as literal maps on \(M^\times\), i.e. absorption follows the head-row
continuation/reversal pair.

\subsection{The two diagnostic finalists}\label{the-final-two-candidates}

After local path-dependence, the two diagnostic finalists are \(\RP\) and
\(\RC\).

\begin{lemma}
\label{lem:12.1}
For \(\RP\), \[
\operatorname{AbsTrans}_{\RP}=\{C,J\}.
\] For \(\RC\), \[
\operatorname{AbsTrans}_{\RC}=\{C^{-1},C^{-1}J\}.
\]
\end{lemma}

\begin{proof}
Let \(x=(r,c)\in M^\times\).

For \(\RP\), \(h(r_x,r_y)=r_y\). The absorption equation is \[
r_y=c.
\] Hence the two absorbed neighbours of \((r,c)\) are the two
off-diagonal points in row \(c\): \[
(c,r),
\qquad
(c,\overline{rc}).
\] These are precisely \[
J(r,c)=(c,r),
\qquad
C(r,c)=(c,\overline{rc}).
\] So \(\operatorname{AbsTrans}_{\RP}=\{C,J\}\).

For \(\RC\), \(h(r_x,r_y)=\overline{r_xr_y}\). The absorption
equation is \[
\overline{r r_y}=c,
\] so \[
r_y=\overline{rc}.
\] The two absorbed neighbours are therefore \[
(\overline{rc},r),
\qquad
(\overline{rc},c),
\] which are \[
C^{-1}(r,c)=(\overline{rc},r),
\qquad
C^{-1}J(r,c)=(\overline{rc},c).
\] Thus \(\operatorname{AbsTrans}_{\RC}=\{C^{-1},C^{-1}J\}\).
\end{proof}

\subsection{Diagnostic separation of the two local
finalists}\label{final-selector-among-the-two-local-finalists}

\begin{theorem}
\label{thm:12.2}
Among the two diagnostic local finalists, the pure \(C/J\)
directed-edge criterion selects \(\RP\).
\end{theorem}

\begin{proof}
By Lemma 12.1, \(\RP\) realizes the undrifted pair
\(\{C,J\}\). \(\RC\) realizes \(\{C^{-1},C^{-1}J\}\), whose
continuation and reversal are drifted by \(C^{-1}\). The pure criterion
selects the literal undrifted pair. The \texttt{L2/} pure-\(C/J\) survivor and transition-factorization
audit tables record the same comparison.
\end{proof}

\subsection{Compact uniqueness core}\label{compact-uniqueness-core}

\begin{theorem}
\label{thm:12.3}
The four conditions \[
\mathcal H_{acc}=0,
\qquad
H_{diag}=0,
\qquad
\text{row-anchor idempotence},
\qquad
\text{pure }C/J
\] select \(\RP\) uniquely inside \(\Omega'\).
\end{theorem}

\begin{proof}
The first condition gives column-blindness by Theorem~\ref{thm:3.1}.
Hence the cross-row rule has the form \[
h(r,r+1)=r+a,
\qquad
h(r,r+2)=r+b
\] for a unique pair \((a,b)\in S^2\). The second condition gives a constant diagonal by
Theorem~\ref{thm:9.1}, and row-anchor idempotence fixes this constant to
\(d=000\) by Proposition~\ref{prop:10.2}.

The remaining condition is pure \(C/J\). Let \(x=(r,c)\in M^\times\)
and write
\(\delta=c-r\in\{1,2\}\). At \(d=000\), absorption of an off-diagonal
point is governed by the cross-row equation \[
x\cdot y=x
\quad\Longleftrightarrow\quad
h(r,r_y)=c.
\] For the two possible cross rows this becomes \[
r_y=r+1\Rightarrow a=\delta,
\qquad
r_y=r+2\Rightarrow b=\delta.
\]

The pure \(C/J\) criterion says that the two absorbed neighbours of
\(x=(r,c)\) are exactly \[
J(r,c)=(c,r),
\qquad
C(r,c)=(c,\overline{rc}).
\] Both lie in row \(c\). Therefore, when \(\delta=1\), absorption must
occur in row \(r+1=c\), forcing \(a=1\). When \(\delta=2\), absorption
must occur in row \(r+2=c\), forcing \(b=2\). Thus \[
(a,b)=(1,2).
\] This is precisely the \(\RP\) cross-row rule, and \(d=000\) has already
been fixed. Hence the four internal conditions select \(\RP\) uniquely.
\end{proof}

The audit table \texttt{l2\_minimal\_selector\_sets.csv} records the
same compact core as a finite audit of the survivor condition.

\subsection{Diagnostic explanatory selection
chain}\label{full-explanatory-selection-chain}

\begin{proposition}
\label{prop:12.4}
The finite diagnostic chain
\[
\Omega'
\xrightarrow{\mathcal H_{acc}=0,
\ H_{diag}=0}
9\times3
\xrightarrow{a\ne b,
\ d=000}
6\times1
\xrightarrow{\min\operatorname{Assoc}_{000}}
2\times1
\xrightarrow{\text{pure }C/J}
1
\]
has terminal survivor \(\RP\).
\end{proposition}

\begin{proof}
The first arrow is Theorems 3.1 and 9.1. The second
arrow is Propositions 10.1 and 10.2. The third arrow is Proposition
10.3. The final arrow is Theorem 12.2. The audit table
\texttt{l2\_selection\_chain.csv} records the survivor count at each
reduction, and the compact L2 verifier checks the displayed chain.
\end{proof}

This diagnostic chain is logically subsumed by Theorem~\ref{thm:12.3}.
After column-blindness, diagonal compression, and row-anchor idempotence, the
pure \(C/J\) condition itself forces \((a,b)=(1,2)\).  The
\(\min\operatorname{Assoc}_{000}\) arrow records the local path-dependence
phenomenon that exposes the \(\RP\)/\(\RC\) ambiguity before the
undrifted directed-edge criterion separates them.


\part{Linear/operator passport after selection}

This part records the finite and carrier-level structures exposed after the
selection theorem.  The role of the notation is now fixed:
\(\RP\) is the selected right-row projection magma, and \(\RC\) is the
second local finalist used for comparison.  The post-selection layer is a passport for the fixed selected
multiplication: it records the operator, term-function, matrix-carrier,
and directed-edge structures forced by \(\RP\).

The passport is organized as follows.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
layer
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
main invariant
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
support
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
row-owner law & every nonempty term has the row of its leftmost variable & structural induction \\
left translations & rank three, image a row fiber, nine-dimensional span & L3 operator audit \\
composition envelope & \(|\operatorname{Mon}^+\langle L_x\rangle|=51\), split into three row-target components & L3 composition tables \\
endomorphisms & six simultaneous permutations and three diagonal collapses & L3 intertwiner audit \\
row contraction & \(E_{ij}J_3E_{kl}^{\top}=E_{ik}\), matching the cross-row \(\RP\) rule & L3 matrix bridge \\
binary terms & \(|T_2(\RP)|=630\), with owner/same-row/cross-row factorization & L3 term tables \\
carrier bridges & matrix/Lie package and signed projective symplectic lift of pure \(C/J\) & L3 bridge audits \\
\end{longtable}

\section{The selected operation and row-owner law}\label{the-selected-rp-operation}

\subsection{Selected multiplication and linearization conventions}\label{rp-multiplication-and-ground-fields}

The selected multiplication is
\[
(r_1,c_1)\cdot(r_2,c_2)=
\begin{cases}
(r_1,r_2), & r_1\ne r_2,\\[2mm]
(r_1,\overline{c_1c_2}), & r_1=r_2,\ c_1\ne c_2,\\[2mm]
(r_1,r_1), & r_1=r_2,\ c_1=c_2.
\end{cases}
\]
The first case is the right-row projection branch; the second is the fixed
same-row Steiner branch; the third is diagonal anchoring.

The finite magma statements are set-theoretic.  When the carrier is
linearized, let \(k\) be a ground field and let \(k[M]\) be the vector
space with basis \(e_x\), \(x\in M\).  Rank and image statements for
individual finite maps are valid over any ground field.  Linear span
statements are over the chosen \(k\).  Spectral, Jordan, and
fiber-algebra statements involving \(t\pm1\) or idempotents with
coefficients \(1/2\) are stated under
\[
\operatorname{char}k\ne2.
\]
The carrier-level matrix/Lie and signed symplectic bridge statements are
over \(\mathbb R\) unless stated otherwise.

The carrier vector space has the matrix-unit identification
\[
k[M]\cong\operatorname{Mat}_3(k),
\qquad e_{(r,c)}\leftrightarrow E_{rc}.
\]
This identification is used for carrier-level linear structures.  The
magma multiplication remains the finite operation displayed above.

\subsection{Row-owner principle}\label{row-owner-principle}

For a nonempty term \(t(x_1,\ldots,x_n)\), define its \textbf{owner}
\(\operatorname{ow}(t)\) to be the leftmost variable appearing in the term
tree.

\begin{lemma}[Row-owner lemma]
\label{lem:13.1}
For every \(\RP\) term \(t(x_1,\ldots,x_n)\),
\[
row\bigl(t(x_1,\ldots,x_n)\bigr)
=
row\bigl(\operatorname{ow}(t)\bigr).
\]
\end{lemma}

\begin{proof}
The assertion is immediate for a variable.  If \(t=u\cdot v\), then the
left-row output axiom gives \(row(t)=row(u\cdot v)=row(u)\).  The induction
hypothesis identifies this row with the row of the leftmost variable of
\(u\), which is also the leftmost variable of \(u\cdot v\).
\end{proof}

The row-owner lemma is the common source of the row-target product law for
left-translation compositions and the owner coordinate in the binary term
algebra.

\section{Operator passport}\label{left-regular-operators}

For \(x\in M\), write \(L_x(y)=x\cdot y\).

\subsection{Left translations}\label{row-fiber-normal-form}

\begin{proposition}
\label{prop:14.1}
For \(x=(r,c)\),
\[
L_{(r,c)}(s,d)=
\begin{cases}
(r,s), & s\ne r,\\[1mm]
(r,r), & s=r,\ d=c,\\[1mm]
(r,\overline{cd}), & s=r,\ d\ne c.
\end{cases}
\]
Consequently \(L_{(r,c)}(M)=F_r\).
\end{proposition}

\begin{proof}
The three cases are exactly the cross-row projection, diagonal anchoring,
and same-row Steiner branches of \(\RP\).  They produce all three points of
\(F_r\).
\end{proof}

\begin{theorem}
\label{thm:14.2}
For every \(x\in M\),
\[
\operatorname{rank}L_x=3,
\qquad
\operatorname{Im}L_x=F_{row(x)}.
\]
Moreover,
\[
\dim\operatorname{span}\{L_x:x\in M\}=9.
\]
\end{theorem}

\begin{proof}
Proposition~\ref{prop:14.1} shows that each \(L_x\) maps onto the
three-point row fiber \(F_{row(x)}\), hence has rank three after
linearization.  Operators with different target rows have disjoint
codomain support.  For a fixed row, the three restrictions to the row
fiber are linearly independent; in normalized row \(0\) they contain the
three matrices
\[
\begin{pmatrix}1&0&0\\0&0&1\\0&1&0\end{pmatrix},
\qquad
\begin{pmatrix}0&1&1\\0&0&0\\1&0&0\end{pmatrix},
\qquad
\begin{pmatrix}0&1&1\\1&0&0\\0&0&0\end{pmatrix}.
\]
Thus each row contributes dimension \(3\), and the total dimension is
\(9\).  The \texttt{L3/} operator rank and spectrum audit tables record the
nine-row calculation.
\end{proof}

Assume \(\operatorname{char}k\ne2\).

\begin{theorem}
\label{thm:14.3}
If \(x=(r,r)\), then
\[
\mu_{L_x}(t)=t(t-1)(t+1)
\]
and \(L_x\) is semisimple.  If \(x=(r,c)\) with \(r\ne c\), then
\[
\mu_{L_x}(t)=t^2(t-1)(t+1),
\]
with zero-primary Jordan structure \(J_2(0)\oplus J_1(0)^{\oplus5}\).
\end{theorem}

\begin{proof}
Every \(L_{(r,c)}\) maps \(k[M]\) onto \(kF_r\), so the nonzero spectral
part is the restriction to \(kF_r\).  For \(c=r\), this restriction fixes
\((r,r)\) and swaps the two off-diagonal basis vectors.  For \(c\ne r\),
the restriction has eigenvalues \(0,1,-1\), and its one-dimensional kernel
lies inside the image, producing exactly one length-two zero block.  The
operator-spectrum table audits the calculation for all nine translations.
\end{proof}

\subsection{Positive composition envelope}\label{composition-envelope-and-directed-edge-representation}

Let \(\operatorname{Mon}^+\langle L_x\rangle\) be the positive composition
monoid generated by the nine left translations.  For \(r\in S\), let
\(\mathcal W_r\) be the component of maps whose image lies in \(F_r\).

For fixed \(r\), define
\[
\tau_{r,c}(d)=
\begin{cases}
r, & d=c,\\
\overline{cd}, & d\ne c,
\end{cases}
\qquad
\mathcal P_r=\langle \tau_{r,c}:c\in S\rangle.
\]
In normalized row \(0\),
\[
\mathcal P_0=\{012,021,200,100,011,022\}.
\]
For \(\psi\in\mathcal P_r\) and \(c\in S\), set
\[
N_{r,\psi,c}(s,d)=
\begin{cases}
(r,\psi(\tau_{r,c}(d))), & s=r,\\
(r,\psi(s)), & s\ne r,
\end{cases}
\]
and let \(K_{r,t}\) be the constant map to \((r,t)\).

\begin{proposition}[Fixed-row composition normal form]
\label{prop:15.1}
For every \(r\in S\),
\[
\mathcal W_r=
\{N_{r,\psi,c}:\psi\in\mathcal P_r,\ c\in S\}
\cup
\{K_{r,t}:t\in S\}.
\]
There are \(14\) nonconstant maps of the first type and \(3\) constants;
hence \(|\mathcal W_r|=17\).
\end{proposition}

\begin{proof}
A positive composition is owned by its leftmost generator.  If all later
generators have the same target row \(r\), the word has the displayed
\(N_{r,\psi,c}\) form.  If some later generator has target row different
from \(r\), the leftmost row-\(r\) generator sees a cross-row input and
the composition becomes constant.  The normalized finite count
\(18-4=14\) for the nonconstant forms is audited in the composition
envelope table.
\end{proof}

\begin{theorem}
\label{thm:15.2}
The positive monoid generated by the left translations has size
\[
|\operatorname{Mon}^+\langle L_x\rangle|=51.
\]
More precisely,
\[
\operatorname{Mon}^+\langle L_x\rangle
=
\mathcal W_0\sqcup\mathcal W_1\sqcup\mathcal W_2,
\qquad |\mathcal W_r|=17.
\]
Its linear envelope decomposes as
\[
\mathcal E=W_0\oplus W_1\oplus W_2,
\qquad
\dim W_r=9,
\qquad
\dim\mathcal E=27,
\]
and satisfies \(W_rW_s\subseteq W_r\).
\end{theorem}

\begin{proof}
Proposition~\ref{prop:15.1} gives \(|\mathcal W_r|=17\), and the three
target-row components are disjoint.  The row-owner law gives
\(W_rW_s\subseteq W_r\).  The linear dimension claim is the row-wise basis-reduction audit recorded
in the \texttt{L3/} composition-envelope and product-law tables.
\end{proof}

\subsection{Row-fiber algebra and directed-edge representation}\label{the-five-dimensional-row-fiber-algebra}

Assume \(\operatorname{char}k\ne2\).  For each row \(r\), define
\[
A_{5,r}=\langle L_{(r,c)}|_{kF_r}:c\in S\rangle
\subseteq\operatorname{End}(kF_r).
\]

\begin{proposition}
\label{prop:15.3}
The algebra \(A_{5,r}\) has dimension \(5\).  Its center is \(kI\), its
radical is two-dimensional and square-zero, and
\[
A_{5,r}/\operatorname{Rad}(A_{5,r})\cong k^3.
\]
For \(r=0\), one basis is
\[
L_{(0,0)}|_{kF_0},\quad L_{(0,1)}|_{kF_0},\quad L_{(0,2)}|_{kF_0},
\quad I,\quad (L_{(0,1)}|_{kF_0})^2.
\]
\end{proposition}

\begin{proof}
By \(\sigma\)-equivariance it is enough to work in row \(0\).  The five-dimensional multiplication table, center, radical, quotient, and
operator-boundary checks are recorded in the auxiliary fiber-algebra
certificate files listed in \texttt{SUPPORT\_INDEX.txt}.
\end{proof}

\begin{theorem}
\label{thm:15.4}
On the six directed edges,
\[
\mathbb R[M^\times]\cong\mathbf1\oplus\mathrm{sign}\oplus2\,\mathrm{std}.
\]
For \(\RP\), the absorption matrix is
\[
A_{abs}=(P_C+P_J)^\top
\]
and has spectrum
\[
\{2,0,0,0,-1,-1\}.
\]
\end{theorem}

\begin{proof}
The six directed edges carry the permutation representation of the
directed triangle.  The pure absorption transitions of \(\RP\) are exactly
\(C\) and \(J\), so the absorption matrix is \((P_C+P_J)^\top\).  The
transition and spectrum tables audit the finite calculation.
\end{proof}

\section{Endomorphisms, intertwiners, and matrix row-contraction}\label{endomorphisms-intertwiners-and-row-contraction}

For a map \(\rho:S\to S\), write
\(\varphi_\rho(r,c)=(\rho(r),\rho(c))\).  For \(t\in S\), write
\(e_t(r,c)=(t,t)\).

For a row \(r\), set
\[
c\star_r d=
\begin{cases}
r, & c=d,\\
\overline{cd}, & c\ne d.
\end{cases}
\]

\begin{lemma}[row-fiber homomorphisms]
\label{lem:16.1}
If \(\phi:S\to S\) satisfies \(\phi(r)=t\) and
\[
\phi(c\star_r d)=\phi(c)\star_t\phi(d)
\qquad(c,d\in S),
\]
then \(\phi\) is either the constant map to \(t\), or a permutation of
\(S\) sending \(r\) to \(t\).
\end{lemma}

\begin{proof}
After translating to \(r=t=0\), write \(\phi(1)=u\), \(\phi(2)=v\).  The
relations \(0\star_0 1=2\) and \(0\star_0 2=1\) force
\[
(u,v)=(0,0),\quad (1,2),\quad (2,1),
\]
which are respectively the constant solution and the two permutations
fixing \(0\).
\end{proof}

\begin{theorem}
\label{thm:16.2}
The endomorphism monoid of the selected \(\RP\) magma is
\[
\operatorname{End}(M,\cdot)
=
\{\varphi_\rho:\rho\in\operatorname{Sym}(S)\}\cup\{e_0,e_1,e_2\}.
\]
\end{theorem}

\begin{proof}
The idempotents of \(\RP\) are exactly the diagonal points \((r,r)\).
Thus an endomorphism sends row fibers to row fibers over a map
\(\rho:S\to S\), and its restriction to each row fiber is governed by
Lemma~\ref{lem:16.1}.  The cross-row rule forces all nonconstant row
restrictions to be the same permutation \(\rho\).  If \(\rho\) fails to be
injective, the cross-row equations force all affected row restrictions to
collapse, and then the third row collapses as well; the map is one of the
\(e_t\).  Conversely, simultaneous permutations and diagonal collapses
preserve the three branches of the \(\RP\) multiplication.
\end{proof}

For an endomorphism \(\varphi\), let \(\operatorname{Hom}_\varphi\) be the
space of intertwiners satisfying
\[
T L_x=L_{\varphi(x)}T
\qquad(x\in M).
\]
Let \(V=k[M]\), and let
\[
\varepsilon:V\to k,
\qquad
\varepsilon\!\left(\sum_{m\in M}a_m e_m\right)=\sum_{m\in M}a_m
\]
be the total augmentation functional.  For \(r\in S\), set
\[
d_r=e_{(r,r)},
\qquad
 o_r=\sum_{c\ne r}e_{(r,c)},
\qquad
U_r=\operatorname{span}\{d_r,o_r\}.
\]
Here \(u\varepsilon\) denotes the rank-one operator
\(v\mapsto \varepsilon(v)u\).

\begin{proposition}[constant intertwiners]
\label{prop:constant-intertwiners}
For each diagonal collapse \(e_r(x)=(r,r)\),
\[
\operatorname{Hom}_{e_r}
=
\{u\varepsilon:u\in U_r\}
=
\left\{
\left(\alpha e_{(r,r)}+\beta\sum_{c\ne r}e_{(r,c)}\right)\varepsilon:
\alpha,\beta\in k
\right\}.
\]
In particular, \(\dim\operatorname{Hom}_{e_r}=2\).
\end{proposition}

\begin{proof}
Each \(L_x\) is induced by a map \(M\to M\), hence every column of its
matrix has sum one and \(\varepsilon L_x=\varepsilon\).  If
\(u\in\operatorname{Fix}(L_{(r,r)})\), then
\[
(u\varepsilon)L_x=u\varepsilon
\qquad\text{and}\qquad
L_{(r,r)}(u\varepsilon)=(L_{(r,r)}u)\varepsilon=u\varepsilon,
\]
so \(u\varepsilon\in\operatorname{Hom}_{e_r}\).

Conversely, if \(T\in\operatorname{Hom}_{e_r}\), then
\[
T L_x=L_{(r,r)}T=T L_y
\qquad(x,y\in M),
\]
so \(T\) kills
\[
\mathcal D=\operatorname{span}\{(L_x-L_y)v:x,y\in M,\ v\in V\}.
\]
Because \(\varepsilon L_x=\varepsilon\), one has
\(\mathcal D\subseteq\ker\varepsilon\).  The reverse inclusion is finite:
the graph generated by relations \(x\cdot z\sim y\cdot z\), with common
source \(z\), is connected.  The Front-C verifier records an explicit
eight-edge spanning tree, so the differences in \(\mathcal D\) span the
augmentation-zero hyperplane.  Thus \(\mathcal D=\ker\varepsilon\), and
\(T\) factors uniquely as \(T=u\varepsilon\).

Substitution into the intertwining equation gives
\(L_{(r,r)}u=u\).  Finally, \(L_{(r,r)}\) has image in the row fiber
\(F_r=\operatorname{span}\{e_{(r,c)}:c\in S\}\); on this fiber it fixes
\(e_{(r,r)}\) and swaps the two off-diagonal basis vectors.  Therefore
\[
\operatorname{Fix}(L_{(r,r)})
=
\operatorname{span}\left\{e_{(r,r)},\sum_{c\ne r}e_{(r,c)}\right\}=U_r,
\]
which proves the formula.
\end{proof}

For automorphisms, \(\operatorname{Hom}_{\varphi_\rho}=kP_\rho\) has
dimension \(1\).  The Front-C verifier audits the defining linear
constraints: the collapse spaces have dimension \(2\), the six
automorphism spaces have dimension \(1\), and therefore the total audit
dimension is
\[
6\cdot1+3\cdot2=12.
\]

\subsection{Matrix row-contraction bridge}\label{cross-row-row-contraction}

\begin{proposition}
\label{prop:16.3}
Under \(k[M]\cong\operatorname{Mat}_3(k)\), the row-contraction extension
of the cross-row \(\RP\) rule is
\[
\mu_{cross}(A,B)=AJ_3B^\top=(A\mathbf1)(B\mathbf1)^\top.
\]
On matrix units with cross rows,
\[
E_{ij}J_3E_{kl}^\top=E_{ik}.
\]
\end{proposition}

\begin{proof}
Since \(J_3=\mathbf1\mathbf1^\top\),
\[
E_{ij}J_3E_{kl}^\top
=(e_ie_j^\top)(\mathbf1\mathbf1^\top)(e_le_k^\top)
=E_{ik}.
\]
For \(i\ne k\), this is exactly the cross-row \(\RP\) rule
\((i,j)\cdot(k,l)=(i,k)\).  The \texttt{L3/} matrix-bridge audit tables record the basis-level bridge.
\end{proof}

Thus the matrix bridge expresses the same column-blindness selected by
\(\mathcal H_{acc}=0\): cross-row multiplication factors through row sums.
The same-row branch remains the Steiner/diagonal row-fiber rule.

\section{Hidden-continuation operator bridge}\label{hidden-continuation-operator-identity}

The local hidden-continuation identity of Section 7 has an operator
interpretation in the left-regular envelope.  For a normalized local index
\(q\), \(D_C\) compares the two stepwise continuation signatures, while
\(D_L\) compares the two collapsed left translations.

\begin{theorem}[Hidden-continuation operator identity]
\label{thm:17.1}
Let \(g\in\Omega'\) be a completed normal-form magma and let
\(q=(b,t,a,e,f)\in S^5\). Define
\[
xy,\qquad yz,\qquad z_R,\qquad ep_L,\qquad ep_R
\]
from \(g\) as in Theorem~\ref{thm:7.1}. Then
\[
\frac{h_q(g)}2
=
D_C(t,xy,f;b,a,z_R)-D_L(ep_L,ep_R).
\]
For brevity write
\[
D_C(q;g)=D_C(t,xy,f;b,a,z_R),
\qquad
D_L(q;g)=D_L(ep_L,ep_R).
\]

After selection, take \(g=\RP\) and write the normalized triple as
\[
X=(0,a),\qquad Y=(b,e),\qquad Z=(t,f).
\]
Then
\[
XY=(0,xy),\qquad YZ=(b,z_R),
\]
and
\[
(XY)Z=(0,ep_L),\qquad X(YZ)=(0,ep_R).
\]
With the convention \(AB=A\circ B\), the continuation distance
\(D_C(q;\RP)\) is the Hamming-distance contrast between the normalized
signatures of
\[
L_{XY}L_Z
\quad\text{and}\quad
L_XL_{YZ},
\]
whereas \(D_L(q;\RP)\) is the Hamming-distance contrast between
\[
L_{(XY)Z}
\quad\text{and}\quad
L_{X(YZ)}.
\]
As maps, all four operators lie in
\(\mathcal W_{\operatorname{row}(X)}\); after linearization they lie in
\(W_{\operatorname{row}(X)}\).

Define
\[
I_{tot}(g)=6\sum_{q\in S^5}D_C(q;g),
\qquad
B_{tot}(g)=6\sum_{q\in S^5}D_L(q;g).
\]
Thus \(H_{tot}(g)=I_{tot}(g)-B_{tot}(g)\). For \(\RP\),
\[
I_{tot}=8904,
\qquad
B_{tot}=1884,
\qquad
H_{tot}=7020,
\qquad
N_-=0.
\]
The finalist \(\RC\) has the same pure hidden-continuation profile.
\end{theorem}

\begin{proof}
The scalar identity is exactly Theorem~\ref{thm:7.1} divided by \(2\):
\(D_C\) is the sum of the nine continuation disagreement indicators,
\(D_L\) is the sum of the nine collapsed left-translation disagreement
indicators, and \(h_q\) is twice their signed difference.

For \(g=\RP\), the displayed normalized triple identifies the two
continuation signatures with the signatures of \(L_{XY}L_Z\) and
\(L_XL_{YZ}\), under the convention \(AB=A\circ B\).  Likewise the
collapsed signatures are the left translations by the endpoints
\((XY)Z\) and \(X(YZ)\).  The row-target assertion follows from the
composition-envelope law \(W_rW_s\subseteq W_r\).  The \texttt{L3/} hidden-continuation bridge audit records the
\(\RP\)/\(\RC\) comparison and verifies the displayed values.  The global frontier status
is Theorem~\ref{thm:7.2}.
\end{proof}

\section{\texorpdfstring{Binary term-function passport: \(T_2(\RP)\)}{Binary term-function passport: T2(RP)}}\label{binary-term-algebra}

A binary term function of \(\RP\) is a function \(M^2\to M\) realized by a
term in variables \(x,y\).  The set of all such functions is denoted
\(T_2(\RP)\).

By the row-owner lemma, every binary term has an owner variable
\(\operatorname{ow}(f)\in\{x,y\}\).  Order the inputs as
\((o,n)=(\operatorname{owner},\operatorname{nonowner})\).  On the same-row
sector, normalize the common row to \(0\) and define
\[
s_f(i,j)=\operatorname{col}\bigl(f(o=(0,i),n=(0,j))\bigr).
\]
On the cross-row sector, normalize the owner row to \(0\) and the nonowner
row to \(1\); the output column is independent of the nonowner column, so
\[
u_f(i)=\operatorname{col}\bigl(f(o=(0,i),n=(1,j))\bigr).
\]
The opposite cross-row direction is determined by \(u_f^{-}(i)=-u_f(-i)\).

\begin{proposition}[Binary sector reduction]
\label{prop:18.1}
Every binary term function has a normal signature
\[
\Theta(f)=\bigl(\operatorname{ow}(f),s_f,u_f\bigr).
\]
The same-row signature \(s_f:S^2\to S\) and cross unary signature
\(u_f:S\to S\) determine the term function on all row sectors.
\end{proposition}

\begin{proof}
The owner coordinate is Lemma~\ref{lem:13.1}.  Same-row inputs remain in
the common input row.  Cross-row inputs are governed by the owner row and a
unary owner-column signature; the opposite direction is forced by the
\(\RP\)-compatible relabelling exchanging the two nonowner directions.
\end{proof}

The realized same-row signatures form a \(21\)-element set
\(\mathcal S_{21}\).  The realized cross unary signatures form a
\(15\)-element set \(\mathcal U_{15}\).  The candidate normal forms are
\[
\{x,y\}\times\mathcal S_{21}\times\mathcal U_{15}.
\]

\begin{lemma}[Binary term certificate protocol]
\label{lem:18.2}
The binary term theorem is supported by four finite obligations:
T2-SIG exposes \(\mathcal S_{21}\) and \(\mathcal U_{15}\); T2-REAL
realizes every triple; T2-TRANS checks closure of the two factor transition
systems; T2-MULT reconstructs the full \(630\times630\) multiplication
table from the owner and factor transitions.
\end{lemma}

\begin{theorem}
\label{thm:18.3}
The binary term-function algebra of \(\RP\) is finite and satisfies
\[
T_2(\RP)
\cong
\{x,y\}\times\mathcal S_{21}\times\mathcal U_{15}.
\]
Consequently
\[
|T_2(\RP)|=2\cdot21\cdot15=630.
\]
\end{theorem}

\begin{proof}
Proposition~\ref{prop:18.1} gives injectivity of the normal signature.
The realization certificate supplies one realizing term for every triple,
and the finite transition tables reconstruct the product law.  Hence the
cardinality is \(2\cdot21\cdot15\).
\end{proof}

This is the binary layer of the term algebra.  The all-arity clone is left
for the open-problem list.

\section{Carrier-level matrix and Lie companion structures}\label{carrier-level-companion-lie-package}

This section records ordinary matrix and Lie structures on the vector-space
carrier exposed after \(\RP\) selection
\cite{HornJohnson2012,FultonHarris1991,Hall2015}.  The bridge is
carrier-level: it uses
\[
k[M]\cong\operatorname{Mat}_3(k)
\]
and the \(\RP\)-compatible simultaneous symbol-permutation action.  The finite
\(\RP\) multiplication and ordinary matrix multiplication are two bilinear
structures on this nine-dimensional carrier.

In this section the reductive decomposition is used only over fields of
characteristic different from \(3\), and the compact-form statements are over
\(\mathbb R\) or \(\mathbb C\).  Ordinary matrix multiplication gives
\[
\mathfrak{gl}(3,k)=kI\oplus\mathfrak{sl}(3,k),
\qquad
\operatorname{Mat}_3(k)=kI\oplus\mathfrak{sl}(3,k)
\]
as a vector-space decomposition.  After complex compact form, this is the
usual central \(\mathfrak u(1)\) plus \(\mathfrak{su}(3)\) picture.

Over \(\mathbb R\), the compact three-dimensional branch
\[
\mathfrak{so}(3)=\{X\in\mathfrak{sl}(3,\mathbb R):X^\top=-X\}
\]
is stable under simultaneous permutation conjugation.  A split comparator is
\(\mathfrak{sl}(V_{std})\subset\mathfrak{sl}(3,\mathbb R)\), where
\(V_{std}=\{v:v_0+v_1+v_2=0\}\).  Compactness selects the
\(\mathfrak{so}(3)\) branch among these three-dimensional
\(\operatorname{Sym}(S)\)-stable carrier branches.

\begin{theorem}
\label{thm:19.1}
In the real/complex compact-form setting, the carrier identification and
\(\RP\)-compatible \(\operatorname{Sym}(S)\)-action expose the following
carrier-level compact companions:
\[
SU(3),
\qquad
SU(2),
\qquad
U(1).
\]
Here \(SU(3)\) is the compact form attached to the ordinary
\(\mathfrak{sl}(3)\) matrix package, \(SU(2)\) is the double cover of the
selected compact \(SO(3)\) branch, and \(U(1)\) is the compact central
factor attached to \(kI\).  The assertion is a carrier-level companion
statement; it does not assert a canonical direct-product action
\(SU(3)\times SU(2)\times U(1)\) generated by the finite magma.
\end{theorem}

\begin{proof}
The matrix-unit carrier gives \(\operatorname{Mat}_3\), hence the usual
\(\mathfrak{sl}(3)\) compact form and the central circle after complex
compact form.  The \(\operatorname{Sym}(S)\)-stable compact branch
\(\mathfrak{so}(3)\) gives \(SO(3)\), with double cover \(SU(2)\).  This
\(SO(3)\) branch is a selected compact sub-branch of the carrier matrix
algebra, not an independent factor commuting with \(SU(3)\).  The \texttt{L3/} companion-structure bridge audit records the
carrier-level compact types.
\end{proof}

\section{\texorpdfstring{Signed symplectic lift of the pure \(C/J\) geometry}{Signed symplectic lift of the pure C/J geometry}}\label{signed-symplectic-bridge}

This bridge uses the finite pure \(C/J\) directed-edge geometry selected in
Part II.  The continuous statement is a signed projective symplectic
realization of that six-edge geometry and completes the carrier-level bridge
part of the passport \cite{McDuffSalamon2017}.

Let \(V=\mathbb R^6\) have ordered basis
\[
(e_{01},e_{12},e_{20},e_{10},e_{02},e_{21})
\]
and symplectic matrix
\[
\Omega=
\begin{pmatrix}0&I_3\\-I_3&0\end{pmatrix}.
\]
Set
\[
P=\begin{pmatrix}0&0&1\\1&0&0\\0&1&0\end{pmatrix},
\qquad
R=\begin{pmatrix}1&0&0\\0&0&1\\0&1&0\end{pmatrix},
\]
\[
\widehat C=\begin{pmatrix}P&0\\0&P\end{pmatrix},
\qquad
\widehat J_0=\begin{pmatrix}0&R\\R&0\end{pmatrix},
\qquad
\widehat J_s=\begin{pmatrix}0&R\\-R&0\end{pmatrix}.
\]
The unsigned \(\widehat J_0\) induces edge reversal but is
anti-symplectic; the signed \(\widehat J_s\) induces the same projective
edge reversal and is symplectic.

\begin{theorem}
\label{thm:20.1}
The finite pure \(\{C,J\}\) directed-edge geometry admits the signed
projective dihedral lift displayed above.  The unsigned reversal is
anti-symplectic, while the signed reversal is symplectic.  The finite tick
word
\[
C,J,C,C,J
\]
visits all six directed edges and has a signed symplectic realization.
\end{theorem}

\begin{proof}
Direct multiplication gives
\[
\widehat C^T\Omega\widehat C=\Omega,
\qquad
\widehat J_0^T\Omega\widehat J_0=-\Omega,
\qquad
\widehat J_s^T\Omega\widehat J_s=\Omega,
\]
and
\[
\widehat C^3=I,
\qquad
\widehat J_s^2=-I,
\qquad
\widehat J_s\widehat C\widehat J_s^{-1}=\widehat C^{-1}.
\]
Hence the projective classes satisfy the dihedral relations in
\(\operatorname{PSp}(V,\omega)\).  On coordinate lines, \(\widehat C\)
acts as \((01\ 12\ 20)(10\ 02\ 21)\), and \(\widehat J_s\) acts as edge
reversal.  Starting from \(01\), the tick word gives
\[
01\mapsto12\mapsto21\mapsto10\mapsto02\mapsto20.
\]
The signed-lift matrices and tick-word audit are recorded in the
\texttt{L3/} signed-symplectic bridge tables.
\end{proof}

This completes the post-selection passport: the finite operator structure,
binary term structure, matrix row-contraction, carrier-level matrix/Lie bridge,
and signed symplectic lift are all read from the single selected multiplication
\(\RP\) and its pure directed-edge geometry.

\part{Certificates and open bridges}

\section{Finite certificate protocol}\label{finite-certificate-protocol}

The support package is available at
\url{https://github.com/burundai-t/finite-row-graded-magmas-over-f3}.

The repository license split is: manuscript files in \texttt{paper/} are
licensed under CC BY 4.0, while all non-paper support files are licensed
under MIT.

It separates three repository roles:
\[
\texttt{L1/},\texttt{L2/},\texttt{L3/}
\quad
\text{frozen finite reference layers},
\]
\[
\texttt{mathcal\_H/}
\quad
\text{the global }\mathcal H\text{-frontier bundle},
\]
\[
\texttt{README.md},\ \texttt{SUPPORT\_INDEX.txt},\ \texttt{tools/},\
\texttt{VERIFY\_LOGS/}
\quad
\text{the repository interface}.
\]
The frozen \texttt{L1/}, \texttt{L2/}, and \texttt{L3/} layers support the
manuscript-facing Landscape/Selection/Linearization finite checks.  The
\texttt{mathcal\_H/} bundle supports the global \(\mathcal H\)-frontier
theorem of Section 7, and
\texttt{mathcal\_H/h2/verify\_h2\_coverage\_light.py} provides the
static coverage check for the accepted H2 raw artifacts.

The reviewer-facing finite-layer check is

\begin{verbatim}
tools/run_quick_checks.sh
\end{verbatim}

It runs the frozen L1/L2/L3 audit path:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
layer
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
role
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
L1 & finite landscape and associativity reference checks \\
L2 & compact selection and pure \(C/J\) survivor checks \\
L3 & selected-\(\RP\) operator, bridge, term-algebra, and companion-structure checks \\
\end{longtable}

The compact table families are:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
family
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
tables
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
purpose
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
L1 landscape atlas & \texttt{L1/tables/} & associativity distribution,
endpoints, controlled \(\mathcal H\) comparison \\
L2 selection & \texttt{L2/tables/} & selection chain and pure \(C/J\)
survivor separation \\
L3 linearization passport & \texttt{L3/tables/} & operator envelope,
\(T_2\), matrix bridge, signed symplectic lift, auxiliary fiber algebra \\
\end{longtable}

The complete file-by-file index, including the bundle root interface and
H2 light coverage support, is \texttt{SUPPORT\_INDEX.txt}.


\subsection{\texorpdfstring{Global \(\mathcal H\)-frontier certificate handles}{Global H-frontier certificate handles}}\label{global-h-frontier-certificate-handles}

The global \(\mathcal H\)-frontier theorem is checked by the separate
\texttt{mathcal\_H/} bundle.  The manuscript-level implication is
\[
\text{local distance reduction}
+
\text{H1--H4 certificate obligations}
\Longrightarrow
\text{Theorem 7.2}.
\]

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
handle
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
bundle location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
theorem component
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
HF-RED, HF-CERT-IF & bundle theorem interface & local distance
reduction \\
H1 & \texttt{h1/} & unrestricted full-landscape range \\
H2 & \texttt{h2/} & pure-frontier UNSAT certificate \\
H3 & embedded verifier check & \(\RP\)/\(\RC\) pure-frontier
witnesses \\
H4 & \texttt{h4/} & signed-cancellation spike classifier \\
\end{longtable}

The exact repository commands for the bundle root and focused H1--H4
checks are listed in \texttt{SUPPORT\_INDEX.txt}.  The H2 obligation proves
\[
(h_q\ge0\ \forall q\in S^5)\wedge rawH\ge1171
\]
UNSAT on the depth-9 live frontier:
\[
18540/18540\ \mathrm{UNSAT},
\qquad SAT=0,
\qquad UNKNOWN=0.
\]
The light static H2 coverage verifier checks the accepted raw H2 artifact
coverage and integrity for the declared live frontier.  This command is the
artifact-coverage audit for H2:

\begin{verbatim}
cd mathcal_H/h2
PYTHONDONTWRITEBYTECODE=1 python3 verify_h2_coverage_light.py
\end{verbatim}

\subsection{Dependency summary}\label{dependency-summary}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
result family
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
support role
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
compact \(\RP\) uniqueness, Theorem 12.3 & hand finite proof; L2 audit
tables \\
explanatory selection chain, Proposition 12.4 & didactic chain,
logically subsumed by Theorem 12.3; local audits \\
global associativity atlas & finite \texttt{L1/} atlas tables \\
global \(\mathcal H\)-frontier & separate H1--H4 certificate
bundle \\
linearization passport & analytic/classification results with L3 audits \\
binary term algebra & sector reduction plus realization/transition
certificates \\
matrix/Lie and signed symplectic structures & bridge audits \\
\end{longtable}

\section{Open problems and bridges}\label{open-problems-and-bridges}

The following fronts are subsequent work. The selection theorem and the
operator passport above are complete at the level stated in this
manuscript.

\subsection{Closures}\label{closures}

\subsubsection{\texorpdfstring{Functorial \(\RP\)--\(\mathfrak{gl}(3)\)
bridge}{Functorial RP--gl(3) bridge}}\label{functorial-rp-gl3-bridge}

The carrier identification \[
k[M]\cong\operatorname{Mat}_3(k)
\] supports two bilinear operations: \(\RP\) multiplication and ordinary
matrix multiplication. The cross-row contraction \[
\mu_{cross}(A,B)=AJ_3B^\top
\] and the carrier-level compact companions give partial alignments. The
closure question is whether there is a universal or functorial
construction from the \(\RP\) magma to the matrix/Lie package, beyond the
shared carrier.

\subsubsection{\texorpdfstring{Symplectic development of pure
\(C/J\)}{Symplectic development of pure C/J}}\label{symplectic-development-of-pure-cj}

The finite pure \(C/J\) geometry has an explicit signed symplectic lift.
The next problem is to develop the associated continuous symplectic
geometry, including canonical \((q,p)\)-charts, Hamiltonian paths, and
coordinate-free separation of pure reversal from drifted reversal.

\subsubsection{Unified selection principle}\label{unified-selection-principle}

Theorem~\ref{thm:12.3} selects \(\RP\) by the conjunction of four finite
selectors of distinct nature: information-theoretic, entropic,
algebraic, and combinatorial-geometric. It remains open whether these
selectors arise from a single natural, non-tautological
information-theoretic or algebraic functional whose global minimizers on
\(\Omega'\) are exactly the \(\RP\) representative. In such a formulation,
\(\RP\) would appear as a global extremum of one canonical object rather
than as the intersection of four independently stated extrema.

\subsection{Extensions}\label{extensions}

\subsubsection{All-arity term algebra}\label{all-arity-term-algebra}

The binary term algebra satisfies \[
|T_2(\RP)|=630,
\qquad
T_2(\RP)\cong\{x,y\}\times\mathcal S_{21}\times\mathcal U_{15}.
\] The all-arity term algebra remains open: the main question is whether
the binary factorization extends to a general clone structure or is
genuinely binary \cite{Post1941,Szendrei1986,Lau2006}.

\subsubsection{Larger carriers}\label{larger-carriers}

Natural generalizations include analogues over larger carriers
\(S\times S\), especially finite fields or cyclic sets, and
directed-edge \(C/J\) geometries beyond the triangle.

\subsubsection{Relaxing AX2}\label{relaxing-ax2}

The landscape \(\Omega'\) fixes the same-row Steiner axiom. The larger
relaxed landscape asks whether the same-row Steiner rule can itself be
selected by intrinsic criteria.

\subsection{Status summary}\label{status-summary}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Front
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Global \(\mathcal H\)-frontier certificate & closed by Theorem 7.2 and the
\texttt{mathcal\_H/} certificate bundle \\
Functorial \(\RP\)--\(\mathfrak{gl}(3)\) bridge & open \\
Unified selection principle & open \\
Symplectic development of pure \(C/J\) & open \\
All-arity term algebra & open \\
Larger carriers & open \\
Relaxing AX2 & open \\
\end{longtable}

\section{Conclusion}\label{conclusion}

The \(\RP\) magma is selected from the finite normal-form landscape
\(\Omega'\cong S^{21}\) by four internal finite criteria: \[
\mathcal H_{acc}=0,
\qquad
H_{diag}=0,
\qquad
\text{row-anchor idempotence},
\qquad
\text{pure }C/J.
\] The surrounding landscape is mapped by the certified associativity atlas and
the \(\mathcal H\)-frontier atlas, and the selected magma then exposes its
linear/operator passport, binary term algebra, carrier-level compact Lie
companions, and signed symplectic bridge.

Thus the theorem package has the intended three-layer form: \[
\boxed{\text{finite landscape}}
\qquad
\boxed{\text{intrinsic finite selection}}
\qquad
\boxed{\text{linearization of }\RP}.
\]


\appendix
\section{Information-theoretic interpretation of the selected magma}\label{appendix-information-theoretic-interpretation}

This appendix records the information-theoretic reading behind the four
selectors used in Theorem~\ref{thm:12.3} and the six basic properties of
the selected magma \(\RP\).  The proofs and finite classifications are in the
main text; the purpose here is to keep the Landauer and Shannon motivations
visible in the final mathematical form of the paper.

\subsection{Two axes: economy and content}\label{two-axes-economy-and-content}

The selected magma is read along two coupled axes.

The first axis is \emph{economy}.  It records which distinctions the
operation can avoid reading, storing, or drifting.  The hidden-column access
selector
\[
\mathcal H_{acc}=I(C_{out};C_1,C_2\mid R_1,R_2)
\]
implements this axis on the cross-row branch.  Its zero value is exactly
column-blindness: after the ordered row pair is known, the output column is
constant over the hidden input columns.  In the Landauer reading, a column
coordinate that is read and then irreversibly discarded contributes a
thermodynamic bookkeeping cost \cite{Landauer1961,Bennett1973,Bennett1982}.
The condition \(\mathcal H_{acc}=0\) selects the rules whose cross-row output
can be computed through the public row interface.

The second axis is \emph{content}.  It records which finite distinctions the
operation preserves: directed public contact, bracket-sensitive histories,
row-fixed self-action, and literal directed-edge continuation/reversal.  The
selected point is economical because the cross-row channel uses only public
row data, and it is nontrivial because the right public row remains visible
in the output.

The rows are therefore the public interface of the interaction, while columns
are the hidden internal coordinate.  The selected cross-row rule
\[
(r_1,c_1)\cdot(r_2,c_2)=(r_1,r_2),
\qquad r_1\ne r_2,
\]
keeps the left row as owner of the result and writes the right row as the
output column.

\subsection{Shannon diagonal compression}\label{shannon-diagonal-compression}

The diagonal selector is the Shannon entropy of the empirical same-point
output distribution
\[
H_{diag}(d)=-\sum_{s\in S}p_s\log_2 p_s.
\]
Shannon entropy measures the number of bits needed to encode the diagonal
output type \cite{Shannon1948,CoverThomas2006}.  The zero-entropy condition
is the finite compression rule: all same-point diagonal outputs carry one
symbol.  Row-anchor idempotence then chooses the normalized row base,
\[
(r,c)^2=(r,r),
\]
so the compressed self-action returns to the diagonal point of the same row.

Thus \(H_{diag}=0\) is the diagonal economy selector, while row-anchor
idempotence is the semantic stabilization of that compressed sector.

\subsection{Six properties of \texorpdfstring{\(\RP\)}{RP}}\label{six-properties-of-rp}

The following table summarizes the interpretation of the six basic properties
that appear across the selection and linearization layers.

\begin{longtable}{@{}p{0.29\linewidth}p{0.16\linewidth}p{0.47\linewidth}@{}}
\toprule
Property of \(\RP\) & Axis & Physical-theoretical unfolding \\
\midrule
\endhead
\bottomrule
\endlastfoot
\(\mathcal H_{acc}=0\) & economy & The cross-row channel uses public rows alone.  In the Landauer reading, the column-read burden is structurally avoided. \\
\((a,b)=(1,2)\) & content & The right public direction remains visible: \(h(r_1,r_2)=r_2\).  Interaction keeps a finite directional distinction while staying column-blind. \\
Low local \(\operatorname{Assoc}_{000}\) among nontrivial column-blind rules & content & Bracket placement carries finite path information.  The diagnostic local selector exposes the two finalist histories before the directed-edge criterion separates them. \\
Pure undrifted \(C/J\) & economy/content & Directed-edge absorption is literal continuation and reversal, \(\operatorname{AbsTrans}_{\RP}=\{C,J\}\).  At the linear level this finite geometry admits the signed symplectic lift of Section~\ref{signed-symplectic-bridge}. \\
\(H_{diag}=0\) & economy & The same-point diagonal stores one output type.  This is the Shannon compression of the diagonal sector. \\
\((r,c)^2=(r,r)\) & content & Self-application returns to the row-fixed point.  Each row has a stable diagonal anchor for self-action. \\
\end{longtable}

The local associativity entry is a diagnostic part of the explanatory chain,
while the compact uniqueness theorem uses the four-condition core
\[
\mathcal H_{acc}=0,
\qquad
H_{diag}=0,
\qquad
\text{row-anchor idempotence},
\qquad
\text{pure }C/J.
\]
After the first three conditions, pure \(C/J\) already forces the cross-row
anchors \((a,b)=(1,2)\), hence the right-row projection rule.

\subsection{Variational reading}\label{variational-reading}

The four-condition selector can be read as a finite variational principle:
\[
\boxed{
\text{minimize hidden information expenditure}
\quad\text{while preserving directed interaction content}.
}
\]
The economy side is supplied by zero hidden-column access and zero diagonal
entropy.  The content side is supplied by row-anchor idempotence, directed public
row dependence, bracket-sensitive local histories, and pure \(C/J\) geometry.

The linear structures appearing after selection are compatible with this
reading:
\[
\begin{aligned}
\mathcal H_{acc}=0
&\leadsto
\text{column-blind row interface and row-contraction},\\
H_{diag}=0
&\leadsto
\text{diagonal compression},\\
(r,c)^2=(r,r)
&\leadsto
\text{row-fixed self-action},\\
(a,b)=(1,2)
&\leadsto
\text{right-row projection},\\
\text{low local }\operatorname{Assoc}_{000}
&\leadsto
\text{path-sensitive interaction},\\
\text{pure }C/J
&\leadsto
\text{dihedral directed-edge geometry and signed symplectic lift}.
\end{aligned}
\]

This semantic reading explains the post-selection passport: the operator,
matrix, term-function, Lie-compatible, and signed symplectic structures are
linear shadows of the same finite row/column economy and content constraints.


\section*{Note on AI assistance}
\addcontentsline{toc}{section}{Note on AI assistance}

This work was developed with the assistance of large language models.

The author specified the conceptual framework, constraints, and target
properties of the objects under study. Within this framework, AI tools
were used for computational exploration, code drafting, and structuring
intermediate results.

All mathematical definitions, statements, and theorems presented in the
paper were established and independently verified by the author. The
finite claims are backed by exhaustive checks recorded in the
accompanying support package.

\newpage
\begin{thebibliography}{99}
\bibitem{AppelHaken1977} K. Appel and W. Haken, ``Every planar map is four colorable. Part I:
Discharging,'' \emph{Illinois Journal of Mathematics} \textbf{21}
(1977), no. 3, 429--490. doi:10.1215/ijm/1256049011.

\bibitem{AppelHakenKoch1977} K. Appel, W. Haken, and J. Koch, ``Every planar map is four colorable.
Part II: Reducibility,'' \emph{Illinois Journal of Mathematics}
\textbf{21} (1977), no. 3, 491--567. doi:10.1215/ijm/1256049012.

\bibitem{Bennett1973} C. H. Bennett, ``Logical reversibility of computation,'' \emph{IBM
Journal of Research and Development} \textbf{17} (1973), no. 6,
525--532. doi:10.1147/rd.176.0525.

\bibitem{Bennett1982} C. H. Bennett, ``The thermodynamics of computation---a review,''
\emph{International Journal of Theoretical Physics} \textbf{21} (1982),
no. 12, 905--940. doi:10.1007/BF02084158.

\bibitem{Birkhoff1935} G. Birkhoff, ``On the structure of abstract algebras,''
\emph{Proceedings of the Cambridge Philosophical Society} \textbf{31}
(1935), no. 4, 433--454. doi:10.1017/S0305004100013463.

\bibitem{Bruck1958} R. H. Bruck, \emph{A Survey of Binary Systems}, Ergebnisse der
Mathematik und ihrer Grenzgebiete, Neue Folge, Heft 20, Springer,
Berlin, 1958.

\bibitem{BurrisSankappanavar1981} S. Burris and H. P. Sankappanavar, \emph{A Course in Universal Algebra},
Graduate Texts in Mathematics 78, Springer-Verlag, New York, 1981.
doi:10.1007/978-1-4613-8130-3.

\bibitem{Cohn1965} P. M. Cohn, \emph{Universal Algebra}, Harper's Series in Modern
Mathematics, Harper \& Row, New York, 1965.

\bibitem{ColbournDinitz2007} C. J. Colbourn and J. H. Dinitz, eds., \emph{Handbook of Combinatorial
Designs}, 2nd ed., Chapman \& Hall/CRC, Boca Raton, 2007.

\bibitem{CoverThomas2006} T. M. Cover and J. A. Thomas, \emph{Elements of Information Theory}, 2nd
ed., Wiley-Interscience, Hoboken, 2006. doi:10.1002/047174882X.

\bibitem{DenesKeedwell1974} J. Dénes and A. D. Keedwell, \emph{Latin Squares and Their
Applications}, Academic Press, New York, 1974.

\bibitem{FultonHarris1991} W. Fulton and J. Harris, \emph{Representation Theory: A First Course},
Graduate Texts in Mathematics 129, Springer-Verlag, New York, 1991.
doi:10.1007/978-1-4612-0979-9.

\bibitem{Gonthier2008} G. Gonthier, ``Formal proof---the four-color theorem,'' \emph{Notices of
the American Mathematical Society} \textbf{55} (2008), no. 11,
1382--1393.

\bibitem{GuilleminSternberg1990} V. Guillemin and S. Sternberg, \emph{Symplectic Techniques in Physics},
Cambridge University Press, Cambridge, 1990.

\bibitem{Hales2005} T. C. Hales, ``A proof of the Kepler conjecture,'' \emph{Annals of
Mathematics} \textbf{162} (2005), no. 3, 1065--1185.
doi:10.4007/annals.2005.162.1065.

\bibitem{HalesEtAl2017} T. Hales, M. Adams, G. Bauer, T. D. Dang, J. Harrison, L. T. Hoang, C.
Kaliszyk, V. Magron, S. McLaughlin, T. T. Nguyen, Q. T. Nguyen, T.
Nipkow, S. Obua, J. Pleso, J. Rute, A. Solovyev, T. H. A. Ta, N. T.
Tran, T. D. Trieu, J. Urban, K. Vu, and R. Zumkeller, ``A formal proof
of the Kepler conjecture,'' \emph{Forum of Mathematics, Pi} \textbf{5}
(2017), e2. doi:10.1017/fmp.2017.1.

\bibitem{Hall2015} B. C. Hall, \emph{Lie Groups, Lie Algebras, and Representations: An
Elementary Introduction}, 2nd ed., Graduate Texts in Mathematics 222,
Springer, Cham, 2015. doi:10.1007/978-3-319-13467-3.

\bibitem{HornJohnson2012} R. A. Horn and C. R. Johnson, \emph{Matrix Analysis}, 2nd ed., Cambridge
University Press, Cambridge, 2012.

\bibitem{Landauer1961} R. Landauer, ``Irreversibility and heat generation in the computing
process,'' \emph{IBM Journal of Research and Development} \textbf{5}
(1961), no. 3, 183--191. doi:10.1147/rd.53.0183.

\bibitem{Lau2006} D. Lau, \emph{Function Algebras on Finite Sets: A Basic Course on
Many-Valued Logic and Clone Theory}, Springer Monographs in Mathematics,
Springer, Berlin, 2006. doi:10.1007/3-540-36023-9.

\bibitem{McCune1997} W. McCune, ``Solution of the Robbins problem,'' \emph{Journal of
Automated Reasoning} \textbf{19} (1997), no. 3, 263--276.
doi:10.1023/A:1005843212881.

\bibitem{McDuffSalamon2017} D. McDuff and D. Salamon, \emph{Introduction to Symplectic Topology},
3rd ed., Oxford Graduate Texts in Mathematics 27, Oxford University
Press, Oxford, 2017.

\bibitem{Pflugfelder1990} H. O. Pflugfelder, \emph{Quasigroups and Loops: Introduction}, Sigma
Series in Pure Mathematics 7, Heldermann Verlag, Berlin, 1990.

\bibitem{Post1941} E. L. Post, \emph{The Two-Valued Iterative Systems of Mathematical
Logic}, Annals of Mathematics Studies 5, Princeton University Press,
Princeton, 1941.

\bibitem{Rosenberg1970} I. G. Rosenberg, ``Über die funktionale Vollständigkeit in den
mehrwertigen Logiken,'' \emph{Rozpravy Československé Akademie Věd, Řada
matematických a přírodních věd} \textbf{80} (1970), no. 4, 3--93.

\bibitem{Shannon1948} C. E. Shannon, ``A mathematical theory of communication,'' \emph{Bell
System Technical Journal} \textbf{27} (1948), 379--423 and 623--656.
doi:10.1002/j.1538-7305.1948.tb01338.x;
doi:10.1002/j.1538-7305.1948.tb00917.x.

\bibitem{Szendrei1986} Á. Szendrei, \emph{Clones in Universal Algebra}, Séminaire de
mathématiques supérieures 99, Presses de l'Université de Montréal,
Montréal, 1986.
\end{thebibliography}

\end{document}
